### 2014-10-29 src/firmas.tex
Template CValles

\thispagestyle{empty}
\cleardoublepage%
\bigskip
\bigskip
\bigskip
\bigskip
\bigskip
\vspace{1cm}
\begin{center}
\includegraphics[scale=0.6]{fig/logo_usm}\\
\cambiaFont{ppl}{\Large \textsc{Examining Committee}}\\
\bigskip
\end{center}
\vspace{2cm}
\begin{tabular}[c]{lc}
{\textsc Prof. Dr. Luis Salinas} & \begin{picture}(150,5) \line(1,0){150} \end{picture}\\
{Universidad Técnica Federico Santa María} & {Advisor}\\
Chile & \\
& \\
& \\
{\textsc Prof. Dr. } &  \begin{picture}(150,5) \line(1,0){150} \end{picture}\\
{Universidad Técnica Federico Santa María} & {Committee Member} \\
Chile & \\
& \\
& \\
{Prof. Dr. } & \begin{picture}(150,5) \line(1,0){150} \end{picture}\\
{University} & {External Peer Reviewer}\\
{Chile} & \\
\end{tabular}

\cleardoublepage%