### 2015-04-01 src/6_Conclusion.tex

\chapter{Conclusion}

\section{Conclusions of this thesis}

\section{Contributions of this thesis}

\section{Future Work}