### 2015-04-01 src/thesis_title.tex
adding professor

\thispagestyle{empty}
\vspace{1.5in}
\begin{center}
{\textsc{Universidad T\'ecnica Federico Santa Mar\'ia}}\\
{\textsc{Departamento de Inform\'atica}}\\
\medskip
\includegraphics[scale=1]{fig/logo_usm.eps}
\bigskip
\bigskip
\bigskip
\bigskip \\
%Titulo
{\textsc{\Large{{Online learning methods for financial time series forecasting}}}}\\%

%\medskip
%{\textsc{\Large{{based on Coreset}}}}%
\bigskip
\bigskip
\bigskip
\bigskip
\bigskip

{\MakeUppercase{Dissertation}\\
\bigskip
\bigskip
Submitted in partial satisfaction of the requirements\\
for the degree of\\
\bigskip
\bigskip
\MakeUppercase{Doctor en Ingenier\'ia Inform\'atica}\\
\bigskip
\bigskip
by\\
\bigskip
{\textsc{\large{Paola Soledad Arce Az\'ocar}}}\\
\bigskip

\bigskip
Advisor\\
Luis Salinas Carrasco\\
\bigskip
\bigskip
In Valpara\'iso, Chile}
\end{center}

\begin{center}
{\normalsize May 2015.}
\end{center}

\pagebreak