### 2015-04-09 src/6_Conclusion.tex
ordenando propuesta 2

\chapter{Conclusion}

\section{Conclusions of this thesis}

%Model effectiveness is focused on out-of-sample forecast rather than in-sample
%fitting. This criterion allows our proposal AVECM prediction capability to be
%expressed rather than just explaining data history.
%
\section{Contributions of this thesis}

\section{Future Work}