### 2015-04-15 src/5_3_Proposal.tex
agregando tercera propuesta

\chapter{Real-time intraday Forex rates forecasting using a regularized Vector Error Correction model}