### 2016-03-22 articles-conferences/icpram2015/main.tex
agregando repo icpram

\documentclass[a4paper,twoside]{article}
\usepackage{graphicx} 
\usepackage{epstopdf}
\usepackage{epsfig}
\usepackage{subfigure}
\usepackage{calc}
\usepackage{amssymb}
\usepackage{amstext}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{lscape}
\usepackage{multicol}
\usepackage{pslatex}
\usepackage{apalike}
\usepackage{SCITEPRESS}
\usepackage[small]{caption}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{enumerate}
\renewcommand{\algorithmicrequire}{\textbf{Input:}}
\renewcommand{\algorithmicensure}{\textbf{Output:}}
\def\R{{\mathbb R }}
\def\N{{\mathbb N }}
\subfigtopskip=0pt
\subfigcapskip=0pt
\subfigbottomskip=0pt
\begin{document}

\title{An Online Vector Error Correction Model for Exchange Rates Forecasting}

\author{\authorname{Paola Arce\sup{1}, Jonathan Antognini\sup{1}, Werner
Kristjanpoller\sup{2} and Luis Salinas\sup{1,3}}
\affiliation{\sup{1}Departamento de Inform\'atica, Universidad T\'ecnica Federico
Santa Mar\'ia, Valpara\'iso, Chile}
\affiliation{\sup{2}Departamento de Industrias, Universidad T\'ecnica Federico
Santa Mar\'ia, Valpara\'iso, Chile}
\affiliation{\sup{3}CCTVal, Universidad T\'ecnica Federico
Santa Mar\'ia, Valpara\'iso, Chile}
\email{\{paola.arce,luis.salinas,werner.kristjanpoller\}@usm.cl, jantogni@csrg.inf.utfsm.cl}
}

\keywords{Vector Error Correction Model, Online Learning, Financial
Forecasting, Foreign Exchange Market}

\abstract{
Financial time series are known for their non-stationary behaviour. However,
sometimes they exhibit some stationary linear combinations. When this happens,
it is said that those time series are cointegrated.The Vector Error Correction
Model (VECM) is an econometric model which characterizes the joint dynamic
behaviour of a set of cointegrated variables in terms of forces pulling towards
equilibrium. In this study, we propose an Online VEC model (OVECM) which
optimizes how model parameters are obtained using a sliding window of the most
recent data. Our proposal also takes advantage of the long-run relationship
between the time series in order to obtain improved execution times. Our
proposed method is tested using four foreign exchange rates with a frequency of
1-minute, all related to the USD currency base. OVECM is compared with VECM and
ARIMA models in terms of forecasting accuracy and execution times. We show that
OVECM outperforms ARIMA forecasting and enables execution time to
be reduced considerably while maintaining good accuracy levels compared with
VECM.
}

\onecolumn \maketitle \normalsize \vfill

\input{src/intro}
\input{src/background}
\input{src/methodology}
\input{src/results}
\input{src/conclusions}

\vfill
\bibliographystyle{apalike}
{\small
\bibliography{reference}}

\vfill
\end{document}