### 2016-03-22 articles-conferences/icpram2015/slides/src/slides.tex

\section{HFT time series}
\subsection{Intro}

\begin{frame}
    \frametitle{Motivation: Financial Time Series Forecasting}
    \begin{itemize}
        \item High Frequency data
        \item Large number of factors affecting the market
    \end{itemize}
    \begin{center}
    \includegraphics[width=0.8\textwidth]{img/financial_ts}
    \end{center}
\end{frame}

\begin{frame}
    \frametitle{Learning Problems}
    \begin{itemize}
        \item Batch Learning
        \begin{itemize}
            \item Training phase then testing
            \item The model doesn't change in testing phase
        \end{itemize}
        \item Online Learning
        \begin{itemize}
            \item No separate phase
            \item "Learning while doing"
        \end{itemize}
    \end{itemize}
\end{frame}

\begin{frame}
    \frametitle{Batch methods}
    \framesubtitle{Ridge Regression}

    \begin{Definition}
    {\em Ridge regression} (RR) is a batch method which is a
        generalization of the least squares method (LS). 
    \end{Definition}

    \begin{equation*} \min_{\mathbf{w}} J(\mathbf{w}), \quad
    \text{where} \quad J(\mathbf{w}) = \sum_{t=1}^m
    (f(\mathbf{x}_t)-y_t)^2  + \gamma R(\mathbf{w}) \, ,
    \end{equation*}

    where $\gamma$ is a regularization parameter and
    $R(\mathbf{w}) = \| \mathbf{w}\| ^2$.
\end{frame}

\begin{frame}
    \frametitle{Batch methods}
    \framesubtitle{Ridge Regression}

    The RR solution $\mathbf{w}_*$ is well-known:

   \begin{equation*}
   {\color{red}\mathbf{w}_*}= (X^\intercal
   X+\gamma \mathbb{I})^{-1}X^\intercal y  
   \end{equation*}

   which is equivalent to:
   \begin{equation*}
   {\color{red}\mathbf{w}_*}= \Big (\sum_{t=1}^m \mathbf{x}_t \mathbf{x}_t
   ^\intercal + \gamma \mathbb{I})\Big )^{-1}
   \sum_{t=1}^m y_t \mathbf{x}_t
   \end{equation*}

\end{frame}

\begin{frame}
    \frametitle{Batch methods}
    \framesubtitle{Ridge Regression}
    The prediction for a new input $\color{blue}\mathbf{x}_{m+1}$ is given by:    
    \begin{equation*}
        f({\color{blue}\mathbf{x}_{m+1}})={\color{red}\mathbf{w}_*^\intercal}
        \color{blue}\mathbf{x}_{m+1}
    \end{equation*}

    \begin{equation*}
        f({\color{blue}\mathbf{x}_{m+1}}) = \displaystyle
        \overbrace{\underbrace{\Big( \sum_{t=1}^m y_t \mathbf{x}_t \Big
        )^\intercal}_{\color{ForestGreen}b} \underbrace{\Big (
        \sum_{t=1}^m \mathbf{x}_t \mathbf{x}_t  ^\intercal + \gamma
        \mathbb{I}}_{\color{ForestGreen}A}\Big)^{-1}}^{{\color{red}\mathbf{w}_*^\intercal}}  \color{blue}\mathbf{x}_{m+1} 
    \end{equation*}
\end{frame}

\begin{frame}
     \frametitle{RR Algorithm}

    \begin{columns}[T]
    \begin{column}{.6\textwidth}
    {\color{purple}\rule{\linewidth}{4pt}}
    \begin{algorithm}[H] \begin{algorithmic}[1] \REQUIRE $\,$ \\
    $\{\mathbf{x}_1,\dots,\mathbf{x}_m \}$: $m$ input vectors \\
    $\{y_1,\dots,y_m \}$: $m$ targets \\ \ENSURE  $\,$ \\
    $\{f(\mathbf{x}_1),\dots,f(\mathbf{x}_m) \}$: model predictions \\
    \STATE Initialize $\mathbf{A}=\gamma \mathbb{I}$ and $\mathbf{b}=0$
    \FOR { $t = 1$ to $m$ } \STATE read new $\mathbf{x}_t$ \STATE output
    prediction $f(\mathbf{x}_t) =  \mathbf{b}^\intercal
    \mathbf{A}^{-1}\mathbf{x}_t$ \STATE $\mathbf{A} = \mathbf{A} +
    \mathbf{x}_t \mathbf{x}_t^\intercal$ \STATE Read new $y_t$ \STATE
    $\mathbf{b} = \mathbf{b} + y_t \mathbf{x}_t$ \ENDFOR
    \end{algorithmic} \caption{Ridge Regression} \label{alg:RR}
    \end{algorithm}
    \end{column}
    \begin{column}{.3\textwidth}
    {\color{blue}\rule{\linewidth}{4pt}}
        \[ {\color{ForestGreen}\mathbf{b}}=  \sum_{t=1}^\intercal y_t \mathbf{x}_t  \]
    \[{\color{ForestGreen}\mathbf{A}}= \sum_{t=1}^\intercal \mathbf{x}_t \mathbf{x}_t
    ^\intercal + \gamma \mathbb{I} \]

    \end{column}
    \end{columns}
\end{frame}

\begin{frame}
    \frametitle{Online methods}
    \framesubtitle{Examples}
    There are several popular online methods such as:
    \begin{itemize}
    \item Perceptron
    \item Passive-aggressive
    \item Stochastic gradient descent
    \item {\color{red} Aggregating algorithm}
    \item The second order perceptron.
    \end{itemize}
\end{frame}

\begin{frame}
     \frametitle{AAR: The Aggregating Algorithm for Regression}

    The {\color{red}AAR} (proposed by Vovk in 2001) is an application of
    the AA to the problem of regression.

    The prediction formula is given by:

    \begin{equation}
    f({\color{blue}\mathbf{x}_{m+1}}) =
    \displaystyle {\Big( \underbrace{\sum_{t=1}^m y_t \mathbf{x}_t \Big
    )}_{\color{ForestGreen}b}}^\intercal \Big ( \underbrace{\sum_{t=1}^{{\color{red}m+1}} \mathbf{x}_t \mathbf{x}_t
    ^\intercal + \gamma \mathbb{I} \Big
    )}_{\color{ForestGreen}\mathbf{A}}^{-1}
    {\color{blue}\mathbf{x}_{m+1}} \, ,
    \end{equation}

    The only difference is the matriz $\color{ForestGreen}\mathbf{A}$ is updated before
    the prediction is made.
\end{frame}

\begin{frame}
     \frametitle{AAR: The Aggregating Algorithm for Regression}

%    Algorithm~\ref{alg:AAR} shows this procedure based on
%    equation~(\ref{eq:predAAR}). This algorithm can be obtained by
%    swapping lines 4 and 5 from algorithm~\ref{alg:RR}. 
    \begin{algorithm}[H] \begin{algorithmic}[1] \REQUIRE $\,$ \\
    $\{\mathbf{x}_1,\dots,\mathbf{x}_m \}$: $m$ input vectors \\
    $\{y_1,\dots,y_m \}$: $m$ targets \\ \ENSURE  $\,$ \\
    $\{f(\mathbf{x}_1),\dots,f(\mathbf{x}_m) \}$: model predictions \\
    \STATE Initialize $\mathbf{A}=\gamma \mathbb{I}$ and $\mathbf{b}=0$
    \FOR { $t = 1$ to $m$ } \STATE read new $\mathbf{x}_t$ \STATE
    $\color{red}\mathbf{A} = \mathbf{A} + \mathbf{x}_t \mathbf{x}_t^\intercal$ \STATE
    {\color{blue}output prediction $f(\mathbf{x}_t) =  \mathbf{b}^\intercal
    \mathbf{A}^{-1}\mathbf{x}_t$} 
    \STATE Read new $y_t$ \STATE $\mathbf{b}
    = \mathbf{b} + y_t \mathbf{x}_t$ \ENDFOR
    \end{algorithmic} \caption{{\em The aggregating algorithm for
    regression}} \label{alg:AAR} \end{algorithm}

\end{frame}

\section{Model} % (fold)
\label{sec:Method}

\begin{frame}
     \frametitle{Motivation}
    Despite of RR and AAR being very successful methods, they have
    some problems:
    \begin{itemize}
        \item They consider all the data for making predictions, but some time series show a strong
    dependence on the latest information instead of all the data.
        \item A matrix inverse is computed at every step of the
        algorithm without considering the matrices are slightly different.
    \end{itemize}
\end{frame}

\begin{frame}
     \frametitle{Proposal}

    The method proposed consists of a modification of AAR considering a
    {\color{red}sliding window} which contains only the last $L+1$ samples 
    $\{\mathbf{x}_m, \mathbf{x}_{m-1}, \dots, \mathbf{x}_{m-L} \}$. 
    If we define an input matrix with the sliding window data:

    \begin{equation*}
    \mathbf{X}(t) = 
    \begin{bmatrix}
    \mathbf{x}_{t-L}^{\intercal}\\
    \vdots \\ 
    \mathbf{x}_t^{\intercal} 
    \end{bmatrix} 
    , \qquad \text{and a target vector} \qquad 
    {\bf y}(t) =
    \begin{bmatrix}
    y_{t-L} \\
    \vdots \\ 
    y_t 
    \end{bmatrix} 
    \end{equation*}
    The optimal solution at each time $t$ is:

    \begin{equation*}
    {\bf w}(t)_*=({\bf
    X}(t)^\intercal{\bf X}(t)+\gamma \mathbb{I})^{-1}{\bf
    X}(t)^\intercal{\bf y}(t) \, .  
    \end{equation*}

\end{frame}

\begin{frame}
     \frametitle{Algorithm}

    \begin{algorithm}[H] \begin{algorithmic}[1] \REQUIRE $\,$ \\
    $\{\mathbf{x}_1,\dots,\mathbf{x}_m \}$: $m$ input vectors \\
    $\{y_1,\dots,y_m \}$: $m$ targets \\ $L$: sliding window size ($L<m$)
    \\ \ENSURE  $\,$ \\ $\{f(\mathbf{x}_{L+1}),\dots,f(\mathbf{x}_m) \}$:
    model predictions \\ \STATE {\color{blue}Initialize $\mathbf{A}=\displaystyle
    \sum_{t=1}^L \mathbf{x}_t \mathbf{x}_t^\intercal + \gamma \mathbb{I}$
    and $\mathbf{b}=\displaystyle \sum_{t=1}^Ly_t \mathbf{x}_t$} \FOR { $t
    = L+1$ to $m$ } \STATE read new $\mathbf{x}_t$

    \STATE $\color{red}\mathbf{A} =
    \mathbf{A} + \mathbf{x}_t \mathbf{x}_t^\intercal- \mathbf{x}_{t-L-1}
    \mathbf{x}_{t-L-1}^\intercal  $ \STATE output prediction
    $f(\mathbf{x}_t) =  \mathbf{b}^\intercal \mathbf{A}^{-1}\mathbf{x}_t$
    \STATE Read new $y_t$ \STATE $\mathbf{b} = \mathbf{b} + y_t
    \mathbf{x}_t$ \ENDFOR

    \end{algorithmic} \caption{Recursive Ridge Regression}
    \label{alg:SLAAR} \end{algorithm}
\end{frame}

\begin{frame}
     \frametitle{Matrix Inverse} 
   Additionally, it is worth noticing that the matrix $\mathbf{A}$ is
   only slightly different:

   \begin{equation*}
        \mathbf{A} = \mathbf{A} + {\color{red}\mathbf{x}_t \mathbf{x}_t^\intercal -\mathbf{x}_{t-L-1}
       \mathbf{x}_{t-L-1}^\intercal}
   \end{equation*}

   However the inverse matrix $\mathbf{A}^{-1}$ can be computed using
   the inverse matrix previously calculated.
\end{frame}

\begin{frame}
     \frametitle{Sherman Morrison formula}
     \begin{theorem}[Sherman Morrison]
     If $\mathbf{A}$ is a positive definite matrix and its inverse
     matrix is known, then the inverse of the matrix
     $\mathbf{B}=\mathbf{A} + \mathbf{x}\mathbf{x}^\intercal$ can be
     obtained as:
     \begin{equation*}
         \mathbf{B}^{-1} = \mathbf{A}^{-1} - 
         \frac{(\mathbf{A}^{-1}\mathbf{x})(\mathbf{A}^{-1}\mathbf{x})^{\intercal}}
         {1 + \mathbf{x}^{\intercal} \mathbf{A}^{-1} \mathbf{x}}
     \end{equation*}
     \end{theorem}   
     This reduces the inverse matrix computation order from $O(n^3)$ to $O(n^2)$.
\end{frame}

\section{Experiments} % (fold)
\label{sec:Experiments}

\begin{frame}
     \frametitle{Experiments}
     \framesubtitle{Data}

    We used daily stock returns of {\color{blue} 45 stocks} from the
    technology sector dated from the {\color{blue}1st of January 2000
    to the 1st of July 2012}.

    Returns $\{x_t\}_{t=1}^{m-1}$ were
    defined based on stock prices $\{{p_t}\}_{t=1}^{m}$ and they are
    related as:

    \begin{equation*} x_t = \frac{p_{t+1}-p_t}{p_t} \, .
    \end{equation*} 
    \begin{block}{Objective}
    The objective is to build a model of the returns of
    stock $k$ based on returns of other stocks of the same financial
    sector.
    \end{block}

\end{frame}

\begin{frame}
     \frametitle{Data}

    Therefore, every input vector $\mathbf{x}_t$ will have 44 stock
    returns at time $t$ without considering information of stock $k$. The
    target vector $\mathbf{y}$ will be the stocks returns of stock $k$.

\end{frame}

\begin{frame}
     \frametitle{Experiments}
     \begin{itemize}
         \item We compared RR, AAR and our proposal.
         \item Experiments were done using 45 stocks.
         \item The error was calculated considering the {\em mean
         squared error} (MSE) using the last 100 predictions made.
     \end{itemize}

\end{frame}

\begin{frame}
     \frametitle{Results} 
    \begin{figure}[H]
      \centering
        \includegraphics[width=0.78\textwidth]{img/data}
      \caption{SPY returns}
      \label{fig:data}
    \end{figure}
\end{frame}

\section{Conclusions} % (fold)
\label{sec:Conclusions}

\begin{frame}
     \frametitle{Conclusions}
    \begin{itemize}
        \item The experiments show that our method was better in 28 of
45 stocks.
        \item The smaller error in our method was for the SPY stock,
        which is an expected result considering that the SPY value is
        constructed based on the main stocks of the markets.
        \item For SPY, it is shown how AAR and
        our method fit better than RR to the target and also that our
        method is slightly better compared with AAR.
        \item It's also possible to notice that when returns change
        drastically none of the methods fit satisfactorily.
    \end{itemize}
\end{frame}

\begin{frame}
     \frametitle{Conclusions}
    \begin{itemize}
    \item The proposal shows an online version of RR which allows the
    number of calculations required to be reduced and it is therefore a better
    option for financial algorithms.
    \item The tests show that in some cases, a portion of data will
    provide a better model instead of using all data.
    \item The historical dependence should be studied in order to
    obtain more accurate algorithms.
    \end{itemize}
\end{frame}