### 2016-03-22 articles-conferences/icpram2015/slides/src/slides_example.tex

\section{HFT time series}
\subsection{Intro}

\begin{frame}
\frametitle{Motivation: Financial Time Series Forecasting}
\begin{columns}
\begin{column}{.48\textwidth}
\color{purple}\rule{\linewidth}{4pt}
Characteristics
  \begin{itemize}
  \item High Frequency data
  \item Large number of factors affecting the market
  \end{itemize}
    \begin{center}
    \includegraphics[width=0.8\textwidth]{img/financial_ts}
    \end{center}
\end{column}%
\begin{column}{.48\textwidth}
\color{blue}\rule{\linewidth}{4pt}
Requires
  \begin{itemize}
  \item Fast implementations
  \item High Frequency Trading Algorithms 
  \end{itemize}
    \begin{center}
    \includegraphics[width=0.9\textwidth]{img/hft}
    \end{center}
\end{column}
\end{columns}

\end{frame}

\begin{frame}
\frametitle{FFN: Feed Forward Neural Network}

FFNs are widely-used to model problems for learning tasks and especially
used to forecast financial time series.

\begin{itemize}
\item FFNs have three layers: 
\begin{enumerate}[a)]
\item Input layer
\item Hidden layer
\item Output layer
\end{enumerate}
\item Each layer consists of a fixed set of neurons
\item All neurons in a layer are fully connected with the other neurons
of the following layer
\item Each connection has an associated weight related to its
dependence
\end{itemize}

\end{frame}

\begin{frame}
  \frametitle{Feed Forward Neural Network}
%\begin{center}
%\includegraphics[width=0.8\textwidth]{img/simple_feedforward_network}
%\end{center}

  %\input{src/ffn1}
\end{frame}

\begin{frame}
  \frametitle{Feed Forward Neural Network}
%\begin{center}
%\includegraphics[width=0.8\textwidth]{img/simple_feedforward_network}
%\end{center}

  %\input{src/ffn}
\end{frame}
\begin{frame}
    \frametitle{FFN: Training and Testing phases}

The learning process to set the weights and biases is done in a
{\color{red}training phase}, which is computationally expensive and
can be computed offline.

However, the output calculation or {\color{red}testing phase} can not
be done offline and needs to be as fast as possible.

\end{frame}

\section{Model}
\subsection{math}

\begin{frame}
\frametitle{Matrix representation}
\framesubtitle{Input Layer}

\begin{columns}
\begin{column}{.48\textwidth}
The input data can be represented as a pair $(\mathbf{x},y)$, where $\mathbf{x}$ is a vector:
\[
        \mathbf{x}=\begin{bmatrix}
        x_1 \\ x_2 \\ \vdots \\ x_n
        \end{bmatrix}
\]

and $y$ is the target.

\end{column}

\begin{column}{.48\textwidth}
In order to simplify the calculations, an extra row set to 1 is added:
\[
        x=\begin{bmatrix}
        x_1  \\ \vdots \\ x_n \\ 1
        \end{bmatrix}
\]

which is done to include the bias.

\end{column}
\end{columns}
\end{frame}

\begin{frame}
\frametitle{Input-Hidden weights}

The connection between the Input and the Hidden layer can be expressed
in a matricial form:
\[
    {\color{blue}
    \mathbf{W}_{h \times {n+1}}} =
    \left[
      \begin{array}{ccc|c}
        w_{1, 1} & \cdots & w_{1, n} & w_{1, n+1} \\
        \vdots   & \ddots & \vdots   & \vdots     \\
        w_{h, 1} & \cdots & w_{h, n} & w_{h, n+1}
      \end{array}
    \right],
  \]
where $w_{i,j}$ is the connection between the hidden neuron $i$ and
the input neuron $j$, $n$ is the length of the input vector and $h$ is
the number of neurons in the hidden layer.
The last column represents the bias for each hidden neuron.

\end{frame}

\begin{frame}
  \frametitle{Hidden-Output weights}

The weight matrix for the Hidden-Output weights is:

  \[
    {\color{blue}
    \mathbf{\Delta}_{m \times h+1}} =
    \left[
      \begin{array}{ccc|c}
        \delta_{1,1} & \cdots & \delta_{1,h} & \delta_{1,h+1}\\
        \vdots & \ddots &  \vdots & \vdots\\
        \delta_{m,1} & \cdots & \delta_{m,h} &\delta_{m,h+1}
      \end{array}
    \right],
  \]

where $\delta_{i,j}$ is the weight between the output neuron $i$ and
the hidden neuron $j$ and $m$ is the number of output neurons. The
bias is represented in the last column.

\end{frame}

\begin{frame}
  \frametitle{State of hidden neurons}

The input of a hidden neuron $i$ is

\[   
     {\color{blue}[\mathbf{Wx}]_{i}}=
     \sum_{j=1}^{n+1}
      w_{i,j}\,x_j 
  \]

and the output

\[
\eta_{i} = \phi \left(
{\color{blue}[\mathbf{Wx}]_{i}}
    \right)
\]

where $\phi$ is the logistic sigmoid function:
\[
\phi(x) = \frac{1}{1+e^{-x}}
\]

\end{frame}

\begin{frame}
  \frametitle{State of output neurons}

The input of an output neuron $i$ is
\[
    {\color{blue}[\mathbf{\Delta \eta}]_{i}} =
      \sum_{j=1}^{h+1}
      \delta_{i,j}\,\eta_i
  \]

and the output

\[
\gamma_{i} = \phi \left(
    {\color{blue}[\mathbf{\Delta \eta}]_{i}} 
    \right)
\]

which is the FFN output.

\end{frame}

\begin{frame}
  \frametitle{Algorithm}

The algorithm in matricial form for an input vector $\mathbf{x}$ is:

  \begin{align*}
    \mathbf{\eta} &= \mathbf{Wx} & \qquad \text{//Input for the
    Hidden layer }\\
    \mathbf{\eta} &= \mathbf{\phi}(\eta) & \qquad \text{//Applying activation
    function}\\
    \mathbf{\gamma} &= \mathbf{\Delta\eta} & \qquad \text{// Input for the
    Output layer}\\
    \mathbf{\gamma} &= \mathbf{\phi(\gamma)} & \qquad \text{//Applying
    activation function}
  \end{align*}

\end{frame}

%\begin{frame}
%  \frametitle{Interaction diagram}
%%  \includegraphics[width=0.7\textwidth]{img/interaction-diagram.eps}
%\begin{center}
%\includegraphics[width=0.9\textwidth]{img/interaction_diagram}
%\end{center}
%\end{frame}

%\begin{frame}
%  \frametitle{CUBLAS}
  %\lstinputlisting{src/cublas.cu}
%\end{frame}

%\begin{frame}
%  \frametitle{CUDA kernels}
  %\lstinputlisting[basicstyle=\small]{src/kernels.cu}
%\end{frame}

\section{Experiments}
\subsection{sol}
\begin{frame}
  \frametitle{Experiments}
  \begin{itemize}
    \item Problem: $n=h=m=1000,\dots,6000$
    \item Two implementations:
    \begin{itemize}
    \item GPU using CUBLAS (CUDA library for Basic Linear Algebra
    Subprograms)
    \item CPU using MKL (Math kernel library)    
    \end{itemize}
    \item Hardware
    \begin{itemize}
    \item Two 6-core Intel Xeon X5650 2.67GHz
    \item NVIDIA M2050 GPU,448 cores, 1.15Ghz    
    \end{itemize}
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Arquitecture of the solution}
  \begin{center}
  \includegraphics[width=0.7\textwidth]{img/arq1}    
  \end{center}
\end{frame}
\begin{frame}
\frametitle{Results}
    \begin{center}
     \includegraphics[width=0.9\textwidth]{img/data}    
    \end{center}
\end{frame}

\section{Conclusions}
\subsection{conclusion}

\begin{frame}
  \frametitle{Conclusions}
  \begin{itemize}
    \item Activation function time is negligible ($\sim$2\%).
    \item The GPU and CPU computation times are comparable when the
    FFN size is small.
    \item The GPU implementation is recommended for a large number of
    neurons (financial models).
    \item ZeroMQ library allows an efficient communication library for
    high frequency data.
  \end{itemize}
\end{frame}

%\section{Future work}
%\subsection{future}
%
\begin{frame}

    \begin{center}

    {\color{red}\Huge Questions?}
    \end{center}

\end{frame}