### 2016-03-22 articles-conferences/icpram2015/src/background.tex

\section{\uppercase{Background}}
\label{sec:background}
\noindent
\subsection{Integration and Cointegration}
A time series $\mathbf{y}$ is said to be integrated of order $d$ if after
differentiating the variable $d$ times, we get an I(0) process, more precisely:

\[
(1-L)^d \mathbf{y} \sim \text{I(0)} \, ,
\]

\noindent where I(0) is a stationary time series and $L$ is the lag operator:

\[
(1-L)\mathbf{y} = \Delta \mathbf{y}=\mathbf{y}_t  -\mathbf{y}_{t-1} \quad \forall t
\]

Let $\mathbf{y}_t = \{\mathbf{y}^1, \dots, \mathbf{y}^l\}$ be a set of $l$
stationary time series I(1) which are said to be cointegrated if a vector
$\beta=[\beta(1),\dots,\beta(l)]^\top \in \mathbb{R}^l$  exists such that the
time series,

\begin{equation}
 \mathbf{Z}_t:= \beta^\top \mathbf{y}_t = \beta(1) \mathbf{y}^1 + \dots + \beta(l) \mathbf{y}^l \sim
 \text{I(0)}\, .
\end{equation}

In other words, a set of I(1) variables is said to be cointegrated if
a linear combination of them exists which is I(0).

\subsection{Vector Autorregresive Models}\label{sec:varvec}

VECM is a special case of VAR model and both describe the joint behaviour of a
set of variables.

VAR($p$) model is a general framework to describe the behaviour of a set of $l$
endogenous variables as a linear combination of their last $p$ values. These
$l$ variables at time $t$ are represented by the vector $\mathbf{y}_t$ as
follows:

\begin{equation*}
\label{eq:variables}
\mathbf{y}_t = 
\begin{bmatrix} y_{1,t} &
y_{2,t} &
\dots &
y_{l,t}
\end{bmatrix}^\top \, ,
\end{equation*}
\noindent where $y_{j,t}$ corresponds to the time series $j$ evaluated at
time $t$.

The VAR(p) model describes the behaviour of a dependent variable in terms
of its own lagged values and the lags of the others variables in the
system. The model with $p$ lags is formulated as the following:

\begin{equation}
\label{eq:var}
 \mathbf{y}_t = \phi_1 \mathbf{y}_{t-1}  + \dots +   \phi_p\mathbf{y}_{t-p}
+ \mathbf{c} + \mathbf{\epsilon}_t \, ,
\end{equation}

\noindent where $t=p+1, \dots N$,  ${\phi_1,\dots,\phi_p}$ are $l \times l$
matrices of real coefficients,
$\mathbf{\epsilon}_{p+1},\dots,\mathbf{\epsilon}_N$ are error terms,
$\mathbf{c}$ is a constant vector and $N$ is the total number of samples.

The VAR matrix form of equation~(\ref{eq:var}) is:

\begin{equation}\label{eq:vareq}
\mathbf{B} = \mathbf{A} \mathbf{X} + 
\mathbf{E} \, , 
\end{equation}

\noindent where:
\begin{eqnarray*}
\mathbf{A} &=&
\begin{bmatrix}
   \mathbf{y}_{p}     & \dots    & \mathbf{y}_{N-1}\\
   \mathbf{y}_{p-1}   & \dots    & \mathbf{y}_{N-2}\\
   \vdots             & \ddots   & \vdots\\
   \mathbf{y}_{1}     & \dots    & \mathbf{y}_{N-p}\\
   1                  & \dots    & 1 
   \end{bmatrix}^\top \, ,\\
\mathbf{B} &=&
\begin{bmatrix}
\quad \\
\mathbf{y}_{p+1} &
\dots & 
\mathbf{y}_N \\
\quad
\end{bmatrix}^\top  \, ,\\
\mathbf{X}&=&
\left[
\begin{array}{ccccc}
    \quad  & \quad & \quad & \quad \\
    \phi_1   & \cdots & \phi_p & \mathbf{c} \\  
    \quad & \quad & \quad & \quad
    \end{array}
\right]^\top \, ,\\
\mathbf{E} &=& 
\begin{bmatrix}
    \quad \\
    \mathbf{\epsilon}_{p+1}  & 
    \dots                & 
    \mathbf{\epsilon}_N \\
    \quad
\end{bmatrix}^\top \, . 
\end{eqnarray*}

Equation~(\ref{eq:vareq})  can be solved using ordinary least squares estimation.

\subsection{VECM}

VECM is a special form of a VAR model for I(1) variables that are also
cointegrated~\cite{banerjee1993}. It is obtained by replacing the form $\Delta
\mathbf{y}_t = \mathbf{y}_t - \mathbf{y}_{t-1}$ in equation (\ref{eq:var}).
VECM is expressed in terms of differences, has an error correction term and 
the following form:

\begin{equation}
 \label{eq:vec}
 \Delta \mathbf{y}_t = 
 \underbrace{ \Omega\mathbf{y}_{t-1}}_\text{Error correction term} + 
 \sum_{i=1}^{p-1}
\phi_i^* \Delta \mathbf{y}_{t-i}  + \mathbf{c} + \mathbf{\epsilon}_t \quad ,
\end{equation}

\noindent where coefficients matrices $\Omega$ and $\phi_i^*$ are
function of matrices $\phi_i$ (shown in equation (\ref{eq:var})) as follows:

\begin{eqnarray*}
\phi_i^* &: =& -\sum_{j=i+1}^{p} \phi_j \\
\Omega &: =& -(\mathbb{I}-\phi_1-\dots-\phi_p) 
\end{eqnarray*}

The matrix $\Omega$ has the following properties~\cite{johansen1995}:
\begin{itemize}
\item If $\Omega = 0$ there is no cointegration
\item If $rank(\Omega)=l$ i.e full rank, then the time series are not
I(1) but stationary
\item If $rank(\Omega)=r,\quad 0 < r < l$ then, there is cointegration
and the matrix $\Omega$ can be expressed as $\Omega =
\alpha \beta^\top$, where $\alpha$ and $\beta$ are $(l \times r)$
matrices and $rank(\alpha)=rank(\beta)=r$.

The columns of $\beta$ contains the cointegration vectors and the rows of
$\alpha$ correspond with the adjusted vectors. $\beta$ is obtained by Johansen
procedure~\cite{johansen1988} whereas $\alpha$ has to be determined as a
variable in the VECM.

It is worth noticing that the factorization of the matrix $\Omega$ is not
unique since for any $r \times r$ nonsingular matrix $H$ we have:

\begin{eqnarray*}
\alpha \beta^\top &=& \alpha \mathbf{HH^{-1}} \beta^\top\\
&=&(\alpha\mathbf{H})(\beta(\mathbf{H}^{-1})^\top)^\top \\
&=& \alpha^*(\beta^*)^\top
\end{eqnarray*}

\noindent with $\alpha^* = \alpha\mathbf{H}$ and $\beta^* =
\beta(\mathbf{H}^{-1})^\top$.

\end{itemize}

If cointegration exists, then equation (\ref{eq:vec}) can be written as follows:

\begin{equation*}
 \label{eq:vecfull}
 \Delta \mathbf{y}_t = \alpha \beta^\top\mathbf{y}_{t-1} 
 + \sum_{i=1}^{p-1} \phi_i^*\Delta
\mathbf{y}_{t-i}  + \mathbf{c} + \mathbf{\epsilon}_t \quad ,
\end{equation*}

\noindent which is a VAR model but for time series differences.

VECM has the same form shown in equation~(\ref{eq:vareq}) but with
different matrices:

\begin{eqnarray}\label{eq:vecmatrix}
 \mathbf{A}&=&
   \begin{bmatrix} 
   \beta^\top \mathbf{y}_{p} & 
   \cdots & \beta^\top \mathbf{y}_{N-1} \\
   \Delta \mathbf{y}_p  & \cdots 
   &\Delta\mathbf{y}_{N-1} \\ 
   \vdots  & \ddots & \vdots \\
   \Delta\mathbf{y}_2  & \cdots 
   & \Delta \mathbf{y}_{N-p+1} \\
   1 & \cdots & 1 
   \end{bmatrix}^\top \label{eq:vecA} \, ,\\
\mathbf{B} & = &
 \begin{bmatrix}
 \quad\\
  \Delta \mathbf{y}_{p+1} & 
  \dots &
  \Delta \mathbf{y}_N \label{eq:vecB}\\
  \quad
 \end{bmatrix}^\top \, ,\\
\mathbf{X}&=&
  \begin{bmatrix}
   \quad \\
   \alpha & \phi_1^* & \cdots & \phi_{p-1}^* & \mathbf{c} \\  
   \quad
   \end{bmatrix}^\top \, ,\\
\mathbf{E} &=&
\begin{bmatrix}
   \quad \\
   \mathbf{\epsilon}_{p+1} &
   \dots &
   \quad &\mathbf{\epsilon}_N \, \\ \quad
\end{bmatrix}^\top 
\end{eqnarray}

VAR and VECM parameters shown in equation (\ref{eq:vareq}) can be solved
using standard regression techniques, such as ordinary least squares (OLS).

\subsection{Ordinary Least Squares method}

When $\mathbf{A}$ is singular, solution to equation~(\ref{eq:vareq}) is given
by the ordinary least squares (OLS) method. OLS consists of minimizing the sum
of squared errors or equivalently minimizing the following expression:

\begin{equation*}
\label{eq:regressionproblem}
\underset{\mathbf{X}}{\text{min}} \quad \| \mathbf{A}\mathbf{\mathbf{X}} - \mathbf{B} \|_2^2
\end{equation*}

\noindent for which the solution $\hat{\mathbf{X}}$ is well-known:

\begin{equation*}
\label{eq:MP}
\hat{\mathbf{X}}=\mathbf{A}^{\!\!+}\,\mathbf{B}
\end{equation*}

\noindent where $\mathbf{A}^{\!\!+}$ is the Moore-Penrose pseudo-inverse
which can be written as follows:

\begin{equation}
\label{eq:pseudoinverse}
\mathbf{A}^{\!\!+}= (\mathbf{A}^{\!\!\top} \mathbf{A})^{-1}\mathbf{A}^{\!\!\top} \, .
\end{equation}

However, when $\mathbf{A}$ is not full rank, i.e
$rank(\mathbf{A})=k <  n \leq m$, $\mathbf{A}^\top \mathbf{A}$ is
always singular and equation~(\ref{eq:pseudoinverse}) cannot be used.
More generally, the pseudo-inverse is best computed using the compact
singular value decomposition (SVD) of $\mathbf{A}$:

\begin{equation*}
    \label{eq:compactsvd}
    \underset{m \times n}{\mathbf{A}}=
    \underset{m \times k}{\mathbf{U_1}} \enskip
    \underset{k \times k}{\Sigma_1} \enskip
    \underset{k \times n}{\mathbf{V}_1^{\top}} \, ,
\end{equation*}

\noindent as follows

\begin{equation*}
\label{eq:pseudoinversesvd}
\mathbf{A}^{\!\!+} = \mathbf{V}_1 \Sigma_1^{-1} \mathbf{U}_1^\top \, .
\end{equation*}