### 2016-03-22 articles-conferences/icpram2015/src/conclusions.tex

\section{\uppercase{Conclusions}}
\label{sec:conclusions}
\noindent A new online vector error correction method was presented. 
We have shown that our proposed OVECM considerably reduces execution times
without compromising solution accuracy.  OVECM was compared with VECM and ARIMA
with the same sliding window sizes and OVECM outperformed both in terms of
execution time. Traditional VECM slightly outperformed our proposal but the
OVECM execution time is lower.
This reduction of execution time is mainly because OVECM avoids the cointegration vector
calculation using the Johansen method. The condition for getting new vectors is
given by the mean\_error variable which controls how many times the Johansen method is
called. Additionally, OVECM introduces matrix optimization in order to get the
new model in an iterative way.
We could see that our algorithm took much less than a minute at every
step. This means that it could also be used with higher frequency data and would
still provide responses before new data arrives.

For future study, it would be interesting to improve the out-of-sample forecast
by considering more explicative variables, to increase window sizes or trying new
conditions to obtain new cointegration vectors.

Since OVECM is an online algorithm which optimizes processing time, it could be
used by investors as an input for strategy robots. Moreover, some technical
analysis methods could be based on its output.