### 2016-03-22 articles-conferences/elsevier-journal-2015/journal.tex
Agregar estos repos a la tesis

\documentclass[12pt,reqno]{amsart}
\usepackage{amsfonts,amssymb,amsmath,amsthm,
            amssymb,amsopn,amsxtra,amstext,amsbsy}
\usepackage{epsfig}
\usepackage[makeroom]{cancel}
\usepackage{array,arydshln}
\usepackage{array}
\usepackage{url}
%\usepackage{tikz}
\usepackage[spanish]{babel}
\usepackage[utf8x]{inputenc}
\usepackage{fancyhdr}
\usepackage{graphicx}
\usepackage{hyperref,wasysym}
\usepackage{color}
\usepackage{enumerate}
\usepackage{algorithm}
\usepackage{algorithmic}
\renewcommand{\algorithmicrequire}{\textbf{Input:}}
\renewcommand{\algorithmicensure}{\textbf{Output:}}

\newtheorem*{myteo}{Teorema}
\newtheorem*{mydef}{Definici\'on}

\voffset=-10mm
\oddsidemargin=0pt
\evensidemargin=0pt
\topmargin=-10mm
\headsep=10mm
\headheight=18pt
\footskip=30pt
\topskip=0mm
\textwidth 167mm
\textheight=230mm
\parindent=5mm
\marginparsep=3mm
\marginparwidth=19mm
\vfuzz2pt % Don't report over-full v-boxes if over-edge is small
\hfuzz6pt % Don't report over-full h-boxes if over-edge is small

%% MYREMARK ENVIRONMENT %%%%%%%%%%%
\newenvironment{mycolor}{\color{red}}{}
\newcounter{myremarkcount}
\setcounter{myremarkcount}{0}
\newsavebox{\myremarkbox}
\newenvironment{myremark}
{% This is the begin code
\medskip
\begin{flushright}
\begin{lrbox}{\myremarkbox}%
\begin{minipage}[t]{0.9\linewidth}
\stepcounter{myremarkcount}
\footnotesize
\begin{mycolor}
{\bf Remark \arabic{myremarkcount}.}
\begin{it}
}
{% This is the end code
\end{it}
\end{mycolor}
\end{minipage}
\end{lrbox}
\fbox{\usebox{\myremarkbox}}
\end{flushright}

\bigskip}

\title{A Time Varying Vector Error Correction Model Applied to Foreign Exchange
Market}
\author{Paola Arce \and Jonathan Antognini \and Luis Salinas \and Werner Kristjanpoller}

\begin{document}
\maketitle

\begin{abstract}
Financial time series are known for their non-stationary behaviour. However,
sometimes they exhibit some stationary linear combinations. When this happens,
it is said that those time series are cointegrated. Vector error correction
model (VECM) is an econometric model which characterizes the joint dynamic
behaviour of a set of cointegrated variables in terms of forces pulling towards
equilibrium.

VECM parameters are obtained using the ordinary least squares (OLS)
method.  Even though OLS is extensively used, it has over-fitting
issues and is computationally expensive. Ridge regression is commonly used
instead of OLS since they include a regularization parameter which could improve
prediction error.

In this study, we propose an online VECM which optimizes how model parameters
are obtained by using only relevant past data and RR instead of OLS.  Our
proposal also takes advantage of the long-run relationship between the time
series in order to obtain improved execution times. Our proposed method is
tested using synthetic data and foreign exchange rates.
Results show that our proposal enables execution time to be reduced by at least
50\% while maintaining good accuracy levels.
\end{abstract}

\input{src/intro}
\input{src/background}
\input{src/methodology}
\input{src/results}
\input{src/conclusions}
\bibliographystyle{apalike}
\bibliography{reference}
%\begin{thebibliography}{9}
%
%\bibitem{smola}
%Bernhard Scholkopf and Alexander J. Smola. 
%\emph{Learning with Kernels: Support Vector Machines, 
%Regularization, Optimization, and Beyond.}.
%MIT Press, Cambridge, MA, USA, 2001
%
%\bibitem{wiki}
%\url{http://en.wikipedia.org/wiki/Reproducing_kernel_Hilbert_space}
%
%\bibitem{wikibooksIsom}
%\url{http://en.wikibooks.org/wiki/Linear_Algebra/Definition_and_Examples_of_Isomorphisms}
%
%\bibitem{wolframREq}
%\url{http://mathworld.wolfram.com/EquivalenceRelation.html}
%
%\bibitem{wikiHomoM}
%\url{http://en.wikipedia.org/wiki/Homomorphism}
%
%
%\end{thebibliography}

\end{document}