### 2016-03-22 articles-conferences/online-vec-article-2014/src/OLS.tex

\section{Regression problems} \label{sec:OLS}

VAR as also VEC model parameters requires solving the following
multiple regression problem:

\begin{equation}
\label{eq:regproblem}
\underset{m \times n}{\mathbf{A}} \enskip \underset{n \times
l}{\mathbf{\phi}} = \underset{m \times l}{\mathbf{B}}
\end{equation}

In the case matrix $\mathbf{X}$ is non singular, its solution is
straight forward:

\begin{equation}
\label{OLSsolution}
    \mathbf{X}=\mathbf{X}^{-1}\mathbf{B}
\end{equation}

\subsection{Ordinary least squares}
When $\mathbf{X}$ is singular, solution to
equation~(\ref{eq:regproblem}) is given by the ordinary least squares
(OLS) method. OLS consists of minimizing the sum of squared errors or
equivalently minimizing the following expression:

\begin{equation}
\label{eq:regressionproblem}
\underset{\mathbf{X}}{\text{min}} \quad \| \mathbf{X}\mathbf{\mathbf{X}} - \mathbf{B} \|_2^2
\end{equation}

\noindent which solution $\hat{\mathbf{X}}$ is well-known:

\begin{equation}
\label{eq:MP}
\hat{\mathbf{X}}=\mathbf{X}^+\mathbf{B}
\end{equation}

\noindent where $\mathbf{X}^+$ is the Moore-Penrose pseudo-inverse
which can be written as follows:

\begin{equation}
\label{eq:pseudoinverse}
\mathbf{X}^+= (\mathbf{X}^\top \mathbf{X})^{-1}\mathbf{X}^\top \, .
\end{equation}

However, when $\mathbf{X}$ is not full rank, i.e
$rank(\mathbf{X})=k <  n \leq m$, $\mathbf{X}^\top \mathbf{X}$ is
always singular and equation~(\ref{eq:pseudoinverse}) cannot be used.
More generally, the pseudo-inverse is best computed using the compact
singular value decomposition (SVD) of $\mathbf{X}$:

\begin{equation}
    \label{eq:compactsvd}
    \underset{m \times n}{\mathbf{X}}=
    \underset{m \times k}{\mathbf{U_1}} \enskip
    \underset{k \times k}{\mathbf{\Sigma_1}} \enskip
    \underset{k \times n}{\mathbf{V_1}^\top}
\end{equation}

\noindent as follows

\begin{equation}
\label{eq:pseudoinversesvd}
\mathbf{X}^+ = \mathbf{V_1\Sigma_1^{-1}U_1^\top}
\end{equation}

\textbf{Demo}\quad

Since the problem shown in equation~(\ref{eq:regproblem}) has not
solution, the minimum norm given by equation~(\ref{OLSsolution}) is
obtained by solving the equivalent problem:

\begin{equation*}
\label{eq:proyectorsol}
\mathbf{X \hat{\phi} = PB} 
\end{equation*}

\noindent where $\mathbf{P=U_1 U_1^\top}$ is the projection onto the
Col($\mathbf{X}$).

Since $\mathbf{V} = [\underset{(n \times k)}{\mathbf{V_1}} |
\underset{(n \times k)}{\mathbf{V_2}}]$ and $\mathbf{V_1^\top V_2 =
0}$ we can express $\mathbf{\hat{\phi}} = \mathbf{V_1 x_1 + V_2 x_2}$
with $\mathbf{x_2=0}$ because $\mathbf{\hat{\phi}}$ lives in the
$\text{Row}(\mathbf{X})$ given by $\mathbf{V_1}$, so we have:

\begin{eqnarray*}
\mathbf{X \hat{\phi}} &=& \mathbf{PB} \\
\mathbf{U_1 \Sigma_1 V_1^\top \hat{\phi}} &=& \mathbf{U_1 U_1^\top B} \\
\mathbf{ V_1^\top \hat{\phi}} &=&  \mathbf{\Sigma_1^{-1} U_1^\top B} \\ 
\mathbf{ V_1^\top V_1 x_1} &=& \mathbf{\Sigma_1^{-1}
U_1^\top B} \\
\mathbf{x_1}&=& \mathbf{\Sigma_1^{-1} U_1^\top B}
\end{eqnarray*}

\noindent from this result we can obtain $\mathbf{\hat{\phi}}$ and
therefore the pseudo-inverse expression:

\begin{eqnarray*}
\mathbf{\hat{\phi}} &=& \mathbf{V_1 x_1} \\
                &=& \mathbf{V_1 \Sigma_1^{-1} U_1^\top B} \\
\mathbf{X^+} &=& \mathbf{V_1 \Sigma_1^{-1} U_1^\top} \, .
\end{eqnarray*}