### 2016-03-22 articles-conferences/online-vec-article-2014/src/data.tex

\section{Foreign Exchange Rate Data} \label{sec:data}

The data used in this article was obtained from Dukascopy historical
bid and ask prices~\cite{dukascopy}. We used currency prices every
minute from the 2nd of January 2012 to the 18th of November 2013,
which were 987454 inputs.

The available currencies pairs were: AUDCAD, AUDCHF, AUDJPY, AUDNZD,
AUDSGD, AUDUSD, CADCHF, CADHKD, CADJPY, CHFJPY, CHFSGD, EURAUD,
EURCAD, EURCHF, EURDKK, EURGBP, EURHKD, EURJPY, EURNOK, EURNZD,
EURPLN, EURSEK, EURSGD, EURTRY, EURUSD, GBPAUD, GBPCAD, GBPCHF,
GBPJPY, GBPNZD, GBPUSD, HKDJPY, NZDCAD, NZDCHF, NZDJPY, NZDUSD,
SGDJPY, USDCAD, USDCHF, USDDKK, USDHKD, USDJPY, USDMXN, USDNOK,
USDPLN, USDRUB, USDSEK, USDSGD, USDTRY, USDZAR, XAGUSD and XAUUSD.