### 2016-03-22 articles-conferences/online-vec-article-2014/src/varvec.tex

\section{Vector autorregresive models}\label{sec:varvec}

VAR as well as VEC are autorregresive models which describe the joint
behaviour of a set variables. Since VEC is a special case of VAR
model, we will formulate the VAR model first and then show how it is
related to the VEC model.

VAR($p$) model is a general framework to describe
the behaviour of a set of $l$ endogenous variables as a linear
combination of their last $p$ values. These $l$ variables at time $t$
are represented by the vector $\mathbf{y}_t$ as follows:

\begin{equation}
\label{eq:variables}
\mathbf{y}_t = 
\begin{bmatrix} y_{1,t} \\
y_{2,t} \\
\vdots \\
y_{l,t}
\end{bmatrix}
\end{equation}
\noindent where $y_{j,t}$ corresponds to the time series $j$ evaluated at
time $t$.

The VAR(p) model describes the behaviour of a dependent variable in terms
of its own lagged values and the lags of the others variables in the
system. The model with $p$ lags is formulated as the following:

\begin{equation}
\label{eq:var}
 \mathbf{y}_t = \phi_1 \mathbf{y}_{t-1}  + \dots +   \phi_p\mathbf{y}_{t-p}
+ \mathbf{c} + \mathbf{\epsilon}_t, \qquad t=p+1, \dots, N
\end{equation}

\noindent where ${\phi_1,\dots,\phi_p}$ are $l \times l$ 
matrices of real coefficients, $\mathbf{\epsilon}_{p+1},\dots,\mathbf{\epsilon}_N$ are
error terms, $\mathbf{c}$ is a constant vector and $N$ is the total number of
samples.

The VAR matrix form is the following:

\begin{equation}
 \label{eq:varmatrix}
               \underbrace{ \begin{bmatrix}
               \quad \\
               \mathbf{y}_{p+1} &
               \mathbf{y}_{p+2} &
               \dots & 
               \mathbf{y}_N \\
               \quad
               \end{bmatrix}}_{\substack{ \mathbf{B}^\top\\l \times (N-p)}}   
= 
                \underbrace{\left[ 
                \begin{array}{ccccc}
                \quad & \quad & \quad & \quad & \quad \\
                \phi_1  & \phi_2 & \cdots & \phi_p & \mathbf{c} \\  
                \quad &\quad & \quad & \quad & \quad
               \end{array} 
               \right]}_{\substack{ \mathbf{X}^\top\\ l \times (l \times p + 1 )}}
\underbrace{\begin{bmatrix}
   \mathbf{y}_{p}  & \mathbf{y}_{p+1} & \dots    & \mathbf{y}_{N-1}\\
   \mathbf{y}_{p-1}  & \mathbf{y}_{p} & \dots    & \mathbf{y}_{N-2}\\
   \vdots        & \vdots   & \ddots   & \vdots\\
   \mathbf{y}_{1} & \mathbf{y}_{2}   & \dots    & \mathbf{y}_{N-p}\\
   1 & 1   & \dots    & 1 
   \end{bmatrix}}_{\substack{ \mathbf{A}^\top\\ (l\times p +1 )\times (N-p)}}
+
\underbrace{\begin{bmatrix}
                \quad \\
              \mathbf{\epsilon}_{p+1}  & 
              \mathbf{\epsilon}_{p+2}  & 
              \dots                & 
              \mathbf{\epsilon}_N \\
              \quad
             \end{bmatrix}}_{\substack{\mathbf{E}^\top\\l \times (N-p) }} 
\end{equation}

\noindent which can be solve using ordinary least squares estimation.

\section{Integration and Cointegration}

A time series $\mathbf{y}$ is called integrated of order $d$ if after
differencing the variable $d$ times, we get a I(0) variable (stationary
process), i.e:
\[
(1-L)^d \mathbf{y} \sim \text{I(0)}
\]
\noindent where I(0) is a stationary time series and $L$ is the lag operator, i.e,
\[
(1-L)\mathbf{y} = \Delta \mathbf{y}
\]
\noindent where $\Delta \mathbf{y}(t) = \mathbf{y}(t)  -\mathbf{y}(t-1) \quad \forall t $.

Let $\mathbf{y}_t = \{\mathbf{y}^1, \dots, \mathbf{y}^l\}$ be a set of $l$ stationary time
series I(1) which are said to be cointegrated if a vector
$\beta=[\beta(1),\dots,\beta(l)]^\top \in \mathbb{R}^l$  exists such that the time series,

\begin{equation}
 \mathbf{Z}_t:= \beta^\top \mathbf{y}_t = \beta(1) \mathbf{y}^1 + \dots + \beta(l) \mathbf{y}^l \sim
 \text{I(0)}
\end{equation}

In words, a set of I(1) variables is said to be cointegrated if
exists a linear combination of them which is I(0).

\subsection{Stationary Time Series}

A strictly stationary times series is one for which the probabilistic behavior
of every collection of values $\{y_{t_1},y_{t_2},\dots,y_{t_L}\}$ is identical
to that of the time shifted set, more precisely: \[ P\{y_{t_1} \leq
c_1,\dots,y_{t_L} \leq c_L\} = P\{y_{t_1+h} \leq c_1,\dots,y_{t_L+h}
\leq c_L\}
\quad \forall L \in \mathbb{N}, \forall h \in \mathbb{Z}\] \noindent where
$c_1,\dots,c_L$ are constants.

This definition is too strong and difficult to assess it from a single data
set. The weak version of this definition imposes conditions only on the two
first two moments.

A weakly stationary time series is a process which mean, variance and auto
covariance do not change over time:

\begin{eqnarray*}
E(Y_t) &=& \mu  \quad \forall t \in \mathbb{N} \\ E(Y^2_t) &=&
\sigma^2  \quad \forall t \in \mathbb{N} \\
\lambda(s,t)&=&\lambda(s+h,t+h) \quad \forall s,t \in \mathbb{N},
\forall h \in \mathbb{Z}
\end{eqnarray*}

\noindent with $\lambda(s,t) = E[(y_s-\mu)(y_t - \mu)]$.

The following example~\cite{johansen1995} helps to illustrate the meaning
of $\beta$:

\textbf{Example:}

If we have two-dimensional process $\mathbf{X}_t$, $t=1,\dots,T$ by:

\begin{eqnarray*}
\mathbf{X}_{1t} &=& \sum_{i=1}^t \epsilon_{1i} + \epsilon_{2t} \\
\mathbf{X}_{2t} &=& a \sum_{i=1}^t \epsilon_{1i} + \epsilon_{3t} 
\end{eqnarray*}

Since $\mathbf{X}_{1t}$ and $\mathbf{X}_{2t}$ are I(1) processes and there
exist a vector $\beta = [a -1]$ such that:

\[
\beta^\top \mathbf{X}_t = a \mathbf{X}_{1t} -\mathbf{X}_{2t} = 
a\epsilon_{2t} - \epsilon_{3t} \sim \text{I(0)}
\]

then, both processes are said to be cointegrated. If we add a I(0) process
$\mathbf{X}_{3t} = \epsilon_{4t}$  we find that there exists two cointegration
vectors now: $\begin{bmatrix}a &-1& 0\end{bmatrix}$ and $\begin{bmatrix}0
&0&1\end{bmatrix}$ since:

\[
\beta^\top \mathbf{X}_t = 
\begin{bmatrix}
a & -1 & 0 \\
0 & 0 & 1
\end{bmatrix} 
\begin{bmatrix} 
\mathbf{X}_{1t} \\
\mathbf{X}_{2t} \\
\mathbf{X}_{3t}
\end{bmatrix} = 
\begin{bmatrix}
a\epsilon_{2t} - \epsilon_{3t} \\
\epsilon_{4t}
\end{bmatrix}
\]

This example shows how cointegration vectors describes the stable relations
between the processes by linear relations that are more stationary than the
original process.

\section{VEC model}

A VEC model is a special form of a VAR model for I(1) variables that
are also cointegrated. The VEC model is obtained by replacing the form
$\Delta \mathbf{y}_t = \mathbf{y}_t - \mathbf{y}_{t-1}$ in equation
(\ref{eq:var}). The VEC model is expressed in terms of differences,
has an error correction term and it has the following form:

\begin{equation}
 \label{eq:vec}
 \Delta \mathbf{y}_t = 
 \underbrace{ \Omega\mathbf{y}_{t-1}}_\text{Error correction term} + 
 \sum_{i=1}^{p-1}
\phi_i^* \Delta \mathbf{y}_{t-i}  + \mathbf{c} + \mathbf{\epsilon}_t \quad ,
\end{equation}

\noindent where coefficients matrices $\Omega$ and $\phi_i^*$ are
function of matrices $\phi_i$ (shown in equation (\ref{eq:var})) as follows:

\begin{eqnarray*}
\phi_i^* &: =& -\sum_{j=i+1}^{p} \phi_j \\
\Omega &: =& -(\mathbb{I}-\phi_1-\dots-\phi_p) 
\end{eqnarray*}

The matrix $\Omega$ has the following properties:
\begin{itemize}
\item If $\Omega = 0$ there is no cointegration
\item If $rank(\Omega)=l$ i.e full rank, then the time series are not
I(1) but stationary
\item If $rank(\Omega)=r,\quad 0 < r < l$ then, there is cointegration
and the matrix $\Omega$ can be expressed as $\Omega =
\alpha \beta^\top$, where $\alpha$ and $\beta$ are $(l \times r)$
matrices and $rank(\alpha)=rank(\beta)=r$.

The columns of $\beta$ contains the cointegration vectors and the rows of
$\alpha$ correspond with the adjusted vectors. $\beta$ is obtained by Johansen
procedure~\cite{johansen1988} whereas $\alpha$ has to be determined as a
variable in the VEC model.

It is worth noticing that the factorization of the matrix $\Omega$ is not unique since for any
$r \times r$ nonsingular matrix $H$ we have:

\begin{eqnarray*}
\alpha \beta^\top &=& \alpha \mathbf{HH^{-1}} \beta^\top\\
&=&(\alpha\mathbf{H})(\beta(\mathbf{H}^{-1})^\top)^\top \\
&=& \alpha^*(\beta^*)^\top
\end{eqnarray*}

\noindent whith $\alpha^* = \alpha\mathbf{H}$ and $\beta^* =
\beta(\mathbf{H}^{-1})^\top$.

Therefore, to obtain unique values, further restrictions on the model
are required.

\end{itemize}
%Johansen  procedure also gives a set of
%eigenvalues ordered in a descending way which do not correspond to
%$\alpha$ but the number of eigenvalues larger than zero determine the rank
%of $\beta$, i.e the number of cointegration relations.

If cointegration exists, then equation (\ref{eq:vec}) can be written as follows:

\begin{equation}
 \label{eq:vecfull}
 \Delta \mathbf{y}_t = \alpha \beta^\top\mathbf{y}_{t-1} 
 + \sum_{i=1}^{p-1} \phi_i^*\Delta
\mathbf{y}_{t-i}  + \mathbf{c} + \mathbf{\epsilon}_t \quad ,
\end{equation}

\noindent which is a VAR model but for time series differences.

The matricial form of the VEC model is the following:

\begin{equation} \label{eq:vecmatrix}
\underbrace{
               \begin{bmatrix}
               \quad\\
                \mathbf{\Delta y}_{p+1} & 
                \dots &
                \mathbf{\Delta y}_N \\
                \quad
               \end{bmatrix}
               }_{\substack{\mathbf{B}^\top\\ l \times (N-p) }} =
   \underbrace{
    \begin{bmatrix}
     \quad \\
     \alpha & \phi_1^* & \cdots & \phi_{p-1}^* & \mathbf{c} \\  
     \quad
     \end{bmatrix} 
     }_{\substack{ \mathbf{X}^\top\\ l \times (l\times p +1)}}
\underbrace{\begin{bmatrix} 
   \beta^\top \mathbf{y}_{p} & 
   \cdots & \beta^\top \mathbf{y}_{N-1} \\
   \mathbf{\Delta y}_p  & \cdots 
   & \mathbf{\Delta y}_{N-1} \\ 
   \vdots  & \ddots & \vdots \\
   \mathbf{\Delta y}_2  & \cdots 
   & \mathbf{\Delta y}_{N-p+1} \\
   1 & \cdots & 1 
   \end{bmatrix}}_{\substack{\mathbf{A}^\top \\ (l \times p +1) \times (N-p) }}
+
\underbrace{\begin{bmatrix}
              \quad \\
              \mathbf{\epsilon}_{p+1} &
              \dots &
              \quad &\mathbf{\epsilon}_N \\ \quad
             \end{bmatrix}}_{\substack{\mathbf{E}^\top\\ l \times (N-p) }} 
\end{equation}

VAR and VEC model parameters shown in equations (\ref{eq:varmatrix}) and
(\ref{eq:vecmatrix}) can be solve using standard regression techniques, such as
ordinary least squares (OLS). However, matrix $\mathbf{A}$ is usually
rank-deficient and OLS solution could not be found. Ridge regression method (RR)
is commonly used instead of OLS when matrices are ill-conditioned or
rank-deficient since it has improved generalization capabilities.