### 2016-03-22 articles-conferences/elsevier-2016/src/conclusions.tex

\section{\uppercase{Conclusions}}
\label{sec:conclusions}
\noindent A new online vector error correction method was presented. The traditional VECM
was modified in order to use it in an online context. We have shown that our
proposed OVECM considerably reduces execution times without compromising
solution accuracy.  However the execution time of OVECM is only better than
SLVECM if we avoid the cointegration vector calculation using the Johansen
method, this is given by the avg\_error variable which controls how many times
the method is called.  We could see that our algorithm took much less than a
minute at every step. This means that it could also be used with higher
frequency data and would still provide responses before new data arrives. 
In all the presented scenarios, MAPE is still very high and for future
study, it would be interesting to improve the out-of-sample forecast by
considering more explicative variables, optimizing the windows size, the
number of lags or trying new condition to obtain new cointegration vectors.
Since OVECM is an online algorithm which optimizes processing time, it could be
used by investors as an input for strategy robots. Moreover, some technical
analysis methods could be based on its output.