### 2016-07-06 articles-conferences/iscef2016/manuscript-computational-economics/main.tex

\RequirePackage{fix-cm}
%
%\documentclass{svjour3}                     
%\documentclass[smallcondensed]{svjour3}     % onecolumn (ditto)
\documentclass[smallextended]{svjour3}       % onecolumn (second format) ******
%\documentclass[twocolumn]{svjour3}          % twocolumn

\smartqed  % flush right qed marks, e.g. at end of proof
\usepackage{natbib}

\usepackage{graphicx}

\usepackage{amsfonts,amssymb,amsmath,
            amssymb,amsopn,amsxtra,amstext,amsbsy}
\usepackage{epsfig}
\usepackage[makeroom]{cancel}
\usepackage{array,arydshln}
\usepackage{multirow,varwidth}
\usepackage{hhline}
\usepackage{url}
\usepackage[utf8x]{inputenc}
\usepackage{fancyhdr}
\usepackage{hyperref,wasysym}
\usepackage{color}
\usepackage{enumerate}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{setspace}
\usepackage{booktabs}
\let\Algorithm\algorithm
\renewcommand\algorithm[1][]{\Algorithm[#1]\setstretch{1.1}}

\renewcommand{\algorithmicrequire}{\textbf{Input:}}
\renewcommand{\algorithmicensure}{\textbf{Output:}}

\usepackage{xcolor}
\usepackage{hyperref}
\usepackage{pmat}

\journalname{Computational Economics}

\begin{document}

\title{Fast and adaptive cointegration based model
for forecasting high frequency financial time series}

\titlerunning{Cointegration based model
for forecasting high frequency financial time series}        % if too long for running head

\author{Paola Arce  \and Jonathan Antognini \and Werner Kristjanpoller \and Luis Salinas }
%\authorrunning{Short form of author list} % if too long for running head

\institute{P. Arce  \at
              Departamento de Inform\'atica, Universidad T\'ecnica Federico Santa
              Mar\'ia\\
              Tel.: +56-32-2654000\\
              \email{paola.arce@usm.cl}           %  \\
%             \emph{Present address:} of F. Author  %  if needed
           \and
           J. Antognini \at
              Departamento de Inform\'atica, Universidad T\'ecnica Federico Santa
              Mar\'ia\\
              \email{jantognini@csrg.cl}             \\
           \and
           W. Kristjanpoller \at
              Departamento de Industrias, Universidad T\'ecnica Federico Santa
              Mar\'ia\\
              %Fax: +123-45-678910\\
              \email{werner.kristjanpoller@usm.cl}             \\
           \and
           L. Salinas \at
              Departamento de Inform\'atica, Universidad T\'ecnica Federico Santa
              Mar\'ia\\
              Centro Cient\'ifico Tecnol\'ogico de Valpara\'iso (CCTVal) \\
              \email{luis.salinas@usm.cl}           %  \\
}
\date{Received: date / Accepted: date}

\maketitle

\begin{abstract}

Cointegration is a long-run property of some non-stationary time series where a
linear combination of those time series is stationary.  This behaviour has been
studied in finance because cointegration restrictions often improve forecasting.
The Vector Error Correction Model (VECM) is a well-known econometric technique
that characterises short-run variations of a set of cointegrated time series
incorporating long-run relationships as an error correction term. VECM has been
broadly used with low frequency time series. We aimed to adapt VECM to be used
in finance with high frequency stream data.

Cointegration relations change in time and therefore VECM parameters must be
updated when new data is available. We studied how forecasting performance is
affected when VECM parameters and the length of historical data used change in
time. We observed that the number of cointegration relationships varies with the
length of historical data used. Moreover, parameters that increased these
relationships in time led to better forecasting performance. Our proposal,
called an Adaptive VECM (AVECM) is to make a parameters grid search that
maximises the number of cointegration relationships in the near past. To ensure
the search can be executed fast enough, we used a distributed environment.

The methodology was tested using four 10-second frequency time series of the
Foreign Exchange market. We compared our proposal with ARIMA and the naive
forecast of the random walk model. Numerical experiments showed that on average
AVECM performed better than ARIMA and random walk. Additionally, AVECM
significantly improved execution times with respect to its serial version.

\keywords{VECM \and Cointegration \and Forex \and MPI \and Parallel algorithm}

\end{abstract}

\input{body}

% BibTeX users please use one of
\bibliographystyle{spbasic}      % basic style, author-year citations
%\bibliographystyle{spmpsci}      % mathematics and physical sciences
%\bibliographystyle{spphys}       % APS-like style for physics

%\bibliographystyle{plain}       % APS-like style for physics

\bibliography{reference}   % name your BibTeX data base

\end{document}