### 2016-07-06 articles-conferences/iscef2016/src/methodology.tex

\section{\uppercase{Methodology}} \label{sec:methodology}
\subsection{Motivation} \label{sec:proposal}

Cointegration vectors can be found applying the Johansen method which uses a
sample of the last historical data. However, VECM assumes cointegration vectors
do not change in time.  In fact, the long-run relationships between the time
series might change due to several economic factors that can lead to structural
breaks in the cointegration relationship \citep{gregoryETal1996}.  In order to
show the number of cointegration vectors which depends on the amount $L$ of
historical data and the number of lags $p$ in the VECM, we used a grid search.
We defined a grid of possible values for $L$ and $p$. $L$ goes throughout
$[2,14]$ hours with a step size of 4 hours and $p$  throughout $[1,5]$ with step
size of 1. The first experiment consisted in determining the number of
cointegration vectors for all combinations of $L$ and $p$.  We used four forex
rates: EURUSD, GBPUSD, USDCHF and USDJPY with 10-second frequency. Data started
at 13:00 GMT of the 13th of August 2014, when the New York and London financial
markets opened.

Figure~\ref{fig:hists} shows the distribution of the number of cointegration
vectors given by the Johansen method for different values of $L=[2,6,10,14]$ and
$p=1$ . This procedure was carried out by a sliding window of historical data
moving 1000 times.

\begin{figure}[!h]
  %\vspace{-0.8cm}
  \centering
  \includegraphics[width=0.9\textwidth]{img/Fig1}
  \caption{Histogram of the number of cointegration vectors using $p=1$. Four
  possible values for $L$ are shown (2, 6, 10 and 14) hours (1 hour = 360 data
  points).}
  \label{fig:hists}
\end{figure}

From section \ref{sec:background} we know that $r=0$ means no cointegration and
$r=l$ (we are using four rates, so $l=4$) reveals that no process is I(1) but
stationary.  The interesting cases of cointegration are those where $r$ lies
strictly between $0$ and $4$, i.e. $0<r<4$.

In order to measure the extent of cointegration, we introduce a
{\em percentage of cointegration\/} as following:
\begin{equation} \label{eq:pcoint}
PC = 
\frac{\#\{ it \mid \text{$it$ has $r$ c.v. with $0<r<l$}\}}
     {\#it}\times 100
\end{equation}
where c.v. stands for cointegration vectors and $it$ is the number of iterations.

The goal of our next experiment was to find a relation between this ratio $PC$
and the performance measure MSE (see equation \ref{eq:MSE}). $L$ was defined
between $[700,1500]$ data points with step size 100 and $p$ takes values between
$[1,5]$ with step size 1.

\begin{figure}[ht!]
  \centering
  \includegraphics[width=0.8\textwidth]{img/Fig2}
  \caption{MSE versus the percentage of cointegration considering 1000
  iterations. Best $L$ found was 1060. Below MSE versus $L$ shows a rapidly
  decreasing behaviour, founding minimum at PC = 80\% }
  \label{fig:cointvsmse}
\end{figure}

Figure~\ref{fig:cointvsmse} shows the relation between MSE and $PC$ and $L$.  We
found that better cointegration percentage leads to better performance accuracy
in terms of reducing MSE.

Therefore we proposed to choose $L$ and $p$ in order to maximise the percentage
of cointegration $PC$ in the near past. This process was done every time that
new data was processed. However, this search can be slow if we try different
values for $L$ and $p$ and therefore we distributed this calculation in order to
reduce searching time.

Our proposal was then a modified version of VECM, with parameter $p$ and the
amount of historical data used $L$ obtained at every step. Only the search of
$L$ and $p$ was done in a distributed environment, since this is the most
expensive routine. Our proposal is called Adaptive Vector Error Correction model
(AVECM). AVECM is detailed in the algorithm \ref{alg:AVECM} which summarises
our proposal.

\begin{algorithm}[ht!]
\begin{algorithmic}[1]
\REQUIRE $\,$ \\
$\mathbf{y}$: matrix with $N$ input vectors and $l$ time series\\
$j$: Starting point of testing \\
$it$: Ending point of testing \\
$ps$: list of $p$ values \\
$Ls$: list of $L$ values ($L<N$) \\
$m$: Iterations to determine parameters ($m < N-L$)\\
\ENSURE  $\,$ \\
$\{ \hat{\mathbf{y}}[1],\dots,\hat{\mathbf{y}}[it]\}$: prediction vectors \\
\FOR { $i =j$ to $it$ }
   \STATE $\mathbf{Y} \gets \mathbf{y}[:,i-1]$
    \STATE $L,p \gets
    \texttt{get\_best\_params}(Ls,ps,m,\mathbf{Y})$
        \STATE $model = \text{VECM}(\mathbf{Y},L, p)$
        \STATE $\hat{\mathbf{y}}[i-j] = model.predict()$
\ENDFOR
\end{algorithmic}
\caption{AVECM: Adaptive VECM.}
\label{alg:AVECM}
\end{algorithm}

The input of AVECM is time series prices which are cointegrated. The starting
point of testing is $j$ and the total number of iterations is $it$. We need to
ensure that $j$ is at least the maximum value of $Ls$. $Ls$ and $ps$ are the
possible values for $L$ and $p$.  The function {\bf get\_best\_params} makes
this grid search on the two vector lists $Ls$ and $ps$ and returns the
parameters $L$ and $p$ which maximise the percentage of cointegration $PC$ (see
equation~\ref{eq:pcoint}) for a pre-defined number of iterations $m$. This
function is implemented in a distributed environment, thus ensuring a response
before new data is available.  After $L$ and $p$ parameters are found, VECM is
built and used to forecast the next data point.

\subsection{Model comparison} \label{sec:random}
We compared our proposal, in terms of performance, with the naive forecast of
the random walk model and ARIMA. It is still difficult to outperform the random
walk model for standard econometric forecasting models \citep{lo2011} despite its
simplicity. The random walk model is defined as:

\begin{equation}
\mathbf{y}_t = \mathbf{y}_{t-1} + \epsilon_{t}
\label{rwmodel}
\end{equation}

The naive forecast of the time series difference $\hat{\mathbf{y}}_{t+1}$ for
the random walk model is defined as:
\begin{equation}
\hat{\mathbf{y}}_{t+1} = \mathbf{y}_t + \hat{\epsilon}_{t+1} 
\end{equation}
\noindent where  $\hat{\epsilon}_{t+1} = \epsilon_{t}$.

On the other hand, ARIMA is widely used to forecast returns in finance
\citep{tsay2005}. A process can be modelled as an ARIMA$(p,d,q)$ model if
$\mathbf{x}_t=\Delta^d \mathbf{y}_t $, i.e after differencing $d$ times the time
series $\mathbf{y}_t$,  we get an ARMA$(p,q)$. Since we are modelling returns,
we used $d=1$.

\subsection{Evaluation methods} \label{sec:evaluation}

Forecast performance was evaluated using two different methods which are
frequently used in finance:
\begin{description}
\item
{\bf MSE},  Mean Square Error measures the distance between forecasts
and the true values and large deviations from the true value have a
large impact due to squaring forecast error.
\begin{equation}\label{eq:MSE}
\text{MSE} = 
\frac{\displaystyle \sum_{t=1}^{N} (\mathbf{y}_t-\hat{\mathbf{y}}_t)^2}{N}
\end{equation}
\item {\bf $U$-statistic}, the Theil's $U$-statistic \citep{theil1966} is a unit
free measure obtained as the ratio between the root MSE (RMSE) of a model and
the RMSE of the naive random walk model. A $U$-statistic less than 1 implies the
performance is better than the naive model.
\end{description}