### 2017-08-27 comentarios-tesis/chapter3ok/20170814_Paola_Chapter_3_again.tex
last changes to chapter 3

%% 20170814_Paola_Chapter_3_again.tex

\red{\section{\Large AUGUST 14, 2017 -- Chapter 3 once again!}}

\begin{description}[style=unboxed,leftmargin=0cm,itemsep=3ex]

\myitem{P 36, section 3.2, L 6}
It says:\quad ``...in finding a learning function $f:X\to Y$
which models... etc.''
It is not said here what are $X$ and $Y$; 
further on it is said. 
This is not usual in a scientific text.
Thus, I would suggest a text like this:

\red{``...in finding a learning function $f:X\to Y$, where $X\subseteq\R^m$
and $Y\subset\R$, which models a set of examples $S\subseteq X\times Y$,
called {\em training sets}, which have been drawn randomly and
independently from $X\times Y$ according to \blue{an} unknown joint
distribution function $p(x,y)$ on $X\times Y$:
$$
S=\big\{ (\mathbf{x}_k,y_k)\in X\times Y \mid k=1,\dots,n \big\}
$$
Here, $Y=\R$ in case of regression problems and $Y=\{-1,1\}$ if it is
a classification problem.''}

\blue{However, below there is a deeper discussion of these facts and
I propose you a more comprehensive composition.}

%% \myitem{P 36, section 3.2, L -4}
%% The expression $S\in\mathscr{T}=\mathscr{P}(X\times Y)$ seems
%% unnecessarily complicated.

%% Isn't it the same (and much simpler) to say that the training sets
%% fulfill $S\subseteq X\times Y$ ?

%% The introduction of $\mathscr{T}$ suggests that you want to have the
%% possibility of choosing \red{several} training sets $S$ taken from
%% $\mathscr{T}=\mathscr{P}(X\times Y)$, but this is unnecessary since
%% if you indeed had several training sets, say, $S_\alpha,\;S_\beta,\;\dots$, 
%% you coud always consider a super-training set
%% $S=S_\alpha \cup S_\beta \cup \dots$ and you would achieve the same generality.

%% Thus in the sequel I shall abandone the set $\mathscr{T}$ and shall assume
%% that we have just \red{ONE} training set $S$ and that this is a \red{subset}
%% of $X\times Y$: $S\subseteq X\times Y$.

\myitem{P 36, section 3.2, L -3}
Suddenly the symbol \red{$\mathscr{H}$} appears on the stage!
But it is not defined before or in the same line!
Only after the formula for $\mathcal{A}$ there is a hint about
$\mathscr{H}$.
This writing style is highly non-standard in math!

\myitem{P 36-37, section 3.2. Some remarks trying to make sense of all these!}

\mbox{}\newline
After reading your definitions several times, it seems to me that:

\begin{enumerate}

\item
Which are the references for the many definitions in these pages ?

\item
The most basic data sets you have are $X\subseteq\R^m$ and $Y\subseteq\R$.
You distinguish two cases: $Y=\R$, the case of regression problems;
and $Y=\{-1,1\}$, the case of classification problems.

\item
Then you have a training set $S\subseteq X\times Y$ of the form:
$$
S=\big\{ (\mathbf{x}_k,y_k)\in X\times Y \mid k=1,\dots,n \big\}\,,
$$
where the members of $S$ have been drawn randomly and independently
according to a unknown joint distribution function
$p(\mathbf{x},y)$ on $X\times Y$.

\item
The problem is to find a function $f:X\to Y$ such that:
\begin{itemize}
\item
it \red{models} the training set $S$ in the sense that
$f(\mathbf{x}_k)=y_k$ for all $k=1,\dots,n$; \\
\red{Is this true? If not, then in what sense $f$ models $S$ ?}

\item
it has a \red{predictive capacity} in the sense that $f(\mathbf{x})=y$
for a newly arrived data $(\mathbf{x},y)\in X\times Y\setminus S$.
\end{itemize}
These conditions seem to be too restrictive.
We can weaken them a little by accepting that the function $f:X\to Y$ should:
\begin{itemize}
\item
just \red{closely models} the training set $S$ in the sense that
$f(\mathbf{x}_k)\approx y_k$ with an error $|f(x_k)-y_k|<\varepsilon$ for
a given $\varepsilon>0$ and all $k=1,\dots,n$; \\
\red{Is this true? If not, then in what sense $f$ models $S$ ?}
\item
have an \red{approximate predictive capacity} in the sense that
$f(\mathbf{x})\approx y$ with error $|f(\mathbf{x})-y|<\varepsilon$
for a newly arrived data $(\mathbf{x},y)\in X\times Y\setminus S$.
\end{itemize}
\item 
This function $f:X\to Y$ should be searched after (and hopefully found)
among a set of functions $\mathscr{H}$ from $X$ to $Y$.

The set of functions $\mathscr{H}$ is called \red{hypothesis space}.

Technically speaking, $\mathscr{H}$ is a subset of the set $Y^X$
of \red{all} functions $\varphi:X\to Y$, but this is mathematically too
technical for our purposes in this work.

For the search process it would be convenient to introduce a topology
on $\mathscr{H}$, preferrable a norm topology (like an $L^p$-norm for
instance) and to require topological completeness.
It would be even better if we would require that $\mathscr{H}$ is Banach
space, but this would be perhaps too technical for our goals here.

\item
There is also a learning algorithm $\mathcal{A}$ which allows $f$
to be found in the \red{hypothesis space} $\mathscr{H}$.

We discussed this subject in October last year.
Please see my remarks from that month.

You now interpret $\mathcal{A}$ as an application (function) from
$\mathscr{T}\subseteq\mathscr{P}(X\times Y)$ to $\mathscr{H}$, 
$\mathcal{A}:S\mapsto\mathcal{A}(S)=f$.

\red{Note that you write $\mathscr{T}=\mathscr{P}(X\times Y)$, but
$\mathscr{T}\subseteq\mathscr{P}(X\times Y)$ gives you more freedom.}

I think there is still a conceptual ``darkness'' associated with 
``$S\in\mathscr{T}\subseteq\mathscr{P}(X\times Y)$''.

The problem is \red{to find a learning function $f$}, with the
properties stated above, for some particular (financial) experiment,
but it is one and only one well defined experiment, for instance,
the analysis of a vector valued (financial) time series (this would
be the $\mathbf{x}_k=\left[x_k^1,\dots,x_k^m\right]^T$ for the
(time-) instants $k=1,\dots,n$, but what would represent the 
``$y_k\in\R$''?
But never mind, it will be a real number (possibly $\pm1$) for each $k$.

Now comes the delicate point: for this one and only one experiment
you have:
\begin{itemize}
\item[1.]
\red{Either} one and only one data set 
$S\in\mathscr{T}\subseteq\mathscr{P}(X\times Y)$
and hence your algorithm $\mathcal{A}$ acting on $S$
produces one and only one learning function $f$, $\mathcal{A}(S)=f$,
for the whole experiment.

In this case it doesn't make sense (you could, but it doesn't help too
much!) to consider the algorithm $\mathcal{A}$
as a mapping $\mathcal{A}:\mathscr{T}\to\mathscr{H}$ since $\mathscr{T}$
has only one member, namely, $\mathscr{T}=\big\{S\big\}$.
\item[2.]
\red{Or} several different data sets
$S_\alpha,S_\beta,\dots\in\mathscr{T}\subseteq\mathscr{P}(X\times Y)$
of the same one and only one experiment, and hence your algorithm
$\mathcal{A}$ acting on $S_\alpha$, $S_\beta$, $\dots$ should produce
learning functions $\mathcal{A}(S_\alpha)=f_\alpha$, 
$\mathcal{A}(S_\beta)=f_\beta$, $\dots$

In this case you could have 
$\mathscr{T}=\big\{ S_\alpha,S_\beta,\dots\big\}$
and it make sense to consider the algorithm $\mathcal{A}$ as a mapping
$\mathcal{A}:\mathscr{T}\to\mathscr{H}$:
$$
S_\alpha\mapsto \mathcal{A}(S_\alpha)=f_\alpha\in\mathscr{H}\,,\quad
S_\beta \mapsto \mathcal{A}(S_\beta )=f_\beta \in\mathscr{H}\,,\quad\dots
$$
\end{itemize}

\item
\blue{\bf An example of Case 2. as above, to fix the ideas!}
This would be the case if you consider the data sets 
$S_\alpha,S_\beta,\dots\in\mathscr{T}\subseteq\mathscr{P}(X\times Y)$
as \red{segments} or \red{windows} of a time series
$\big\{(\mathbf{x}_t,y_t)\big\}_{t\in\Z}$ of vectors 
$(\mathbf{x}_t,y_t)\in\R^m\times\R$, say
\begin{align*}
S_\alpha &= \bigg\{ 
\left( \mathbf{x}_{t_\alpha}, y_{t_\alpha} \right), 
\left( \mathbf{x}_{t_\alpha+1}, y_{t_\alpha+1} \right), \dots, 
\left( \mathbf{x}_{t_\alpha+n-1}, y_{t_\alpha+n-1} \right) \bigg\}\,, \\
S_\beta &= \bigg\{ 
\left( \mathbf{x}_{t_\beta}, y_{t_\beta} \right), 
\left( \mathbf{x}_{t_\beta+1}, y_{t_\beta+1} \right),\dots, 
\left( \mathbf{x}_{t_\beta+n-1}, y_{t_\beta+n-1} \right) \bigg\}\,, \\
\vdots & \rule{10em}{0ex}\vdots
\end{align*}
for $t_\alpha,t_\beta,\dots\in\Z$, which without restriction we can assume
to be increasing: $t_\alpha<t_\beta<\dots$.

Note that here all data sets $S_\alpha,S_\beta,\dots$ have the same cardinality
(windows of the same length), namely $n$.
But this is not obligatory: windows could have different length, but
computations would then be more cumbersome.

Also note that nothing is said about the convergence of the sequence
$f_\alpha,f_\beta,\dots$
In fact this sequence may very well not converge in $\mathscr{H}$.
The terms of this sequence are functions that model the behaviour of the
time series within the particular window associated to each data set
$S_\alpha,S_\beta,\dots$ and what is expected is that the corresponding
learning functions $f_\alpha,f_\beta,\dots$ have good predictive capacity,
namely that: 
\begin{align*}
\widehat{y}_{t_\beta}
&:= \mathcal{A}(S_\alpha)(\mathbf{x}_{t_\beta})
\equiv f_\alpha(\mathbf{x}_{t_\beta}) \approx y_{t_\beta} \quad
\text{with an error $\big|\widehat{y}_{t_\beta}-y_{t_\beta}\big|<\varepsilon$}\,, \\
\widehat{y}_{t_\gamma}
&:= \mathcal{A}(S_\beta)(\mathbf{x}_{t_\gamma})
\equiv f_\beta(\mathbf{x}_{t_\gamma}) \approx y_{t_\gamma} \quad
\text{with an error $\big|\widehat{y}_{t_\gamma}-y_{t_\gamma}\big|<\varepsilon$}\,, \\
\vdots & \rule{10em}{0ex}\vdots
\end{align*}
Note that in these equations $f_\alpha$ is evaluated in $\mathbf{x}_{t_\beta}=$
the $X$-coordinate of the first member of the next ``window'' $S_\beta$.

Similarly, $f_\beta$ is evaluated in $\mathbf{x}_{t_\gamma}=$ the $X$-coordinate
of the first member of the next ``window'' $S_\gamma$.
Etc.

Note, however, that each data set in Case 2. is just another version of
the one and only one data set in Case 1.
Thus, Case 2. is \red{not} a generalisation of Case 1.
In fact, Case 2. can be reduced to Case 1. just by taking the union of
all data sets of Case 2.

\red{\bf Thus, it seems to me that you are considering Case 2.
I will assume this in the sequel!}

\item\label{referenceFunction}
Note that there is always an {\em observed\/} or {\em reference function\/}
$$
\mathscr{Y}:X\to Y\,,\qquad
\mathscr{Y}: \mathbf{x}_k \mapsto \mathscr{Y}(\mathbf{x}_k):=y_k\,,
\quad k=1,\dots,n\,.
$$
We will always assume that $\mathscr{Y}\in\mathscr{H}$.
In fact, $\mathscr{Y}$ is from the very beginning the most natural member
of $\mathscr{H}$!

Besides this reference function $\mathscr{Y}$ there are the learning
functions
$f\in\mathscr{H}$:
$$
f:X\to Y\,,\qquad
f: \mathbf{x}_k \mapsto f(\mathbf{x}_k)=:\widehat{y}_k\,,
\quad k=1,\dots,n\,.
$$
The {\em error function\/} associated to any $f\in\mathscr{H}$
is defined by:
$$
\texttt{Error}_f(\mathbf{x}_k)
:=|f(\mathbf{x_k})-\mathscr{Y}(\mathbf{x}_k)|=|\widehat{y}_k-y_k|\,,
\quad k=1,\dots,n\,,
$$
where the error function \texttt{Error} should ideally be small, say, 
$\texttt{Error}(\mathbf{x}_k)<\varepsilon$ for all $k=1,\dots,n$, for
a given $\varepsilon>0$.

Note that $f,\mathscr{Y}:X\to Y$ but $\texttt{Error}_f:X\to\R_0^+$.

In function-analytic notation we observe that, for $f\in\mathscr{H}$,
$\texttt{Error}_f$ is a function on $X$:
$$
\texttt{Error}_f:X\to\R_0^+\,,\qquad
\texttt{Error}_f(\mathbf{x}) 
= |f(\mathbf{x})-\mathscr{Y}(\mathbf{x})| 
= |\widehat{y} - y| \quad\forall x\in X\,.
$$
and since the functions $f$, $\mathscr{Y}$, and hence $\texttt{Error}_f$,
are defined on the whole of $X$ we can also write for any
$f\in\mathscr{H}$:
$$
\texttt{Error}_f(\mathbf{x})
= |f(\mathbf{x})-\mathscr{Y}(\mathbf{x})| 
= |f(\mathbf{x}) - y| = |\widehat{y} - y|\,,\quad \mathbf{x}\in X,\ 
\text{provided that}\ (\mathbf{x},y)\in S\in\mathscr{T}.
$$

\item\label{lossFunction}
For a given $S\in\mathscr{T}$, the algorithm $\mathcal{A}$ selects
one (and only one) learning function $f\in\mathscr{H}$ by minimising
the expected error of a {\em loss function\/}:
$$
V:Y\times Y\to\R_0^+\,,\qquad V: (y_1,y_2)\mapsto V(y_1,y_2)\,.
$$
Note that both arguments of $V$ are points in $Y$.
The prototype of this loss function in the case $Y=\R$ is the {\em metric\/}
\begin{equation}\label{BasicLossFunction}
V(y_1,y_2):=|y_1-y_2|\qquad \forall y_1,y_2\in Y=\R\,. 
\end{equation} 
Two simple loss functions for regression problemas derived from
(\ref{BasicLossFunction}) for $f\in\mathscr{H}$ are:
$$
V\big( f(\mathbf{x}),y \big) 
= |f(\mathbf{x})-y| = |\widehat{y}-y|\,,\quad 
\mathbf{x}\in X\ \text{and}\ y\in Y\ 
\text{with}\ (\mathbf{x},y)\in S\in\mathscr{T}\,,
$$
which is the so called $L^1$-{\em loss function\/} and:
$$
V\big( f(\mathbf{x}),y \big) 
= \big(f(\mathbf{x})-y\big)^2 = \big( \widehat{y}-y\big)^2\,,\quad 
\mathbf{x}\in X\ \text{and}\ y\in Y\ 
\text{with}\ (\mathbf{x},y)\in S\in\mathscr{T}\,,
$$
which is the so called {\em square\/} or $L^2$-{\em loss function\/}.

\item
Note that $V\big( f(\mathbf{x}),y \big)$, $\mathbf{x}\in X$ and $y\in Y$ 
with $(\mathbf{x},y)\in S\in\mathscr{T}$, is actually a {\em function\/}
on $X\times Y$:
$$
V\big(f(\mathbf{\cdot}),\cdot):X\times Y\to\R_0^+\,,\quad
\big( \mathbf{x},y \big) \mapsto V\big( f(\mathbf{x}),y \big)\,,\ 
\text{where}\ (\mathbf{x},y)\in S\in\mathscr{T}\,.
$$
Condition ``$(\mathbf{x},y)\in S\in\mathscr{T}$'' is a small nuisance here,
since the function ``$V\big(f(\mathbf{\cdot}),\cdot)$''  \red{is not defined
{\em for all\/} $\mathbf{x}\in X$ and {\em all\/} $y\in Y$.}
But this is inevitable, since the members of the data sets which we are
dealing with are of the form $(\mathbf{x},y)$.
It could be even the case that the variable $\mathbf{x}$ does not run on
the whole of $X$, and this is most surely the case for $y$ regarding $Y$.
When we come to compute the expected value of the function
$V\big(f(\mathbf{\cdot}),\cdot)$, we will see an easy solution to
this ``beauty defect'' by means of a clever probability distribution on
$X\times Y$ (see (\ref{expectedError}) below ).

There is another function closely related to $V\big(f(\mathbf{\cdot}),\cdot)$
but different fro that one:
$$
V\big(f(\cdot), \mathscr{Y}(\cdot)\big):X\to Y\,,\qquad
\mathbf{x}\mapsto V\big(f(\mathbf{x}), \mathscr{Y}(\mathbf{x})\big)
= \big( \widehat{y}, y \big)\,,\qquad f\in\mathscr{H}\,.
$$
Note that here we do not need the ``ugly'' condition
``$(\mathbf{x},y)\in S\in\mathscr{T}$'' because this is built-in in the
definition of the reference function $\mathscr{Y}:X\to Y$
(see (\ref{referenceFunction}) above).

\item\label{expectedError}
Consider a {\em learning function\/} $f\in\mathscr{H}$,
$f(\mathbf{x})=\widehat{y}$.
Also consider the {\em reference function\/} $\mathscr{Y}:X\to Y$, 
$\mathscr{Y}(\mathbf{x})=y$, where $(\mathbf{x},y)\in S\in\mathscr{T}$,
$S\subseteq X\times Y$.

Note that $S$ may be considered as the {\em graph\/} of $\mathscr{Y}$:
$$
\text{Graph}(\mathscr{Y})
:= \big\{ (\mathbf{x}, y)\in X\times Y \mid \mathscr{Y}(\mathbf{x})=y\big\}
= S\subseteq X\times Y\quad !
$$
There are clearly (many) probability distributions $p$ on $X\times Y$
such that:
\begin{equation}\label{probdist}
p\ \text{assigns probability 0 to all ``events''}\  
(\mathbf{x},y)\in X\times Y\setminus\text{Graph}(\mathscr{Y})
= X\times Y\setminus S\,.
\end{equation}
Consider, in addition, a given {\em loss function\/} $V:Y\times Y\to\R_0^+$.
(see (\ref{lossFunction}) above) and a probability distribution $p$ over
$X\times Y$ with property (\ref{probdist}).

The {\em expected error\/} of $V(f(\cdot),\cdot)$ may then be defined as:
$$
E\big[ V( f(\cdot),\cdot) \big]
= \int_{X\times Y} V\big( f(\mathbf{x},y \big)\,dp(\mathbf{x},y)\,.
$$
\color{red}{
Compare this last formula with your formula (3.3).
Note that in your formula it doesn't make sense to have the variables
$\mathbf{x}$ and $y$ on the left-hand-side since this variebles are
used in the integration on the righ-hand-side, hence they are 
``destroyed'' by the integration and they are not available any more:
the whole formul IS NOT a function of $\mathbf{x}$ and $y$.

Also, in these formulas adopt the convention of writing the variable
$\mathbf{x}$ in {\bf boldface} to recall that this variable is a
vector in $X=\R^m$.

}

%%%%%%% DRAFT PROPOSAL %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\vspace{4ex}
\item
{\color{blue}
{\Large\bf A draft proposal.} \\[1ex]
The basic setting for supervised learning is a {\em data set\/} $X$, 
and an  {\em outcome set\/} $Y$.
For our purposes $X$ will be $\R^m$, $m\in\N$, and $Y$ either $\R$
in the case of regression problems, or $\{-1,1\}$ in the case of
classification problems.
In addition we have a plurality of {\em training sets\/}
$$
S=\bigg\{ (\mathbf{x}_k,y_k)\in X\times Y \mid k=1,\dots,n \bigg\}
\subseteq X\times Y\,,
$$
where the members of $S$ have been drawn randomly and independently
from $X\times Y$ according to an {\em unknown\/} joint distribution
function $p(\mathbf{x},y)$ on $X\times Y$.
We gather all these training sets in a subset $\mathscr{T}$ of
$\mathscr{P}(X\times Y)$.

{\em Supervised learning\/} consists then in finding a
{\em learning function\/} $f:X\to Y$ such that:

%% \begin{itemize}
%% \item
%% it \red{models} the training set $S$ in the sense that $f(x_k)=y_k$
%% for all $k=1,\dots,n$ and
%% \item
%% it has a \red{predictive capacity} in the sense that $f(x)=y$ for a
%% newly arrived data $(x,y)\not\in S$.
%% \end{itemize}
%% These conditions seem to be too restrictive.
%% We can weaken them a little by accepting that the function $f:X\to Y$ should:

\begin{itemize}
\item
it closely models the training set $S$ in the sense that
$f(\mathbf{x}_k)\approx y_k$ with an error
$|f(\mathbf{x}_k)-y_k|<\varepsilon$ for a given $\varepsilon>0$
and all $k=1,\dots,n$;
\red{Is this true? If not, then in what sense $f$ models $S$ ?}
\item
it has an approximate predictive capacity in the sense that
$f(\mathbf{x})\approx y$ with error $|f(\mathbf{x})-y|<\varepsilon$
for a newly arrived data $(\mathbf{x},y)\in X\times Y\setminus S$.
\end{itemize}
The learning function $f:X\to Y$ is to be searched after among a set
of functions $\mathscr{H}$ from $X$ to $Y$.
The set of functions $\mathscr{H}$ is called {\em hypothesis space\/}.

%% For the search process it is convenient to consider a vector space
%% structure and a topology on the hypothesis space $\mathscr{H}$.
%% A popular topology in the field is the $L^p$-norm topology for some
%% $p\in\R$ with $1\leq p\leq \infty$.
%% Topological completeness of $\mathscr{H}$ is convenient as well.
%% A convenient comprehensive requirement would be that $\mathscr{H}$ be
%% a Banach space.

In addition, there is a {\em learning algorithm\/} $\mathcal{A}$ that
associates to every data set one and only one learning function
$f\in\mathscr{H}$.
Thus $\mathcal{A}$ can be considered as a function:
$$
\mathcal{A}:\mathscr{T}\to\mathscr{H}\,,\qquad
S\mapsto\mathcal{A}(S)=f\quad\forall S\in\mathscr{T}\,.
$$

PLEASE CONTINUE HERE WITH THE PERTINENT REMARKS WRITTEN ABOVE !

}

\end{enumerate}

\end{description}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\red{\section{\Large AUGUST 19, 2017 -- Chapter 3 once again!}}

\def\bfx{\mathbf{x}}

\begin{description}[style=unboxed,leftmargin=0cm,itemsep=3ex]

\myitem{P 37, L 3}
It says: ``\dots selection of $f$ is based\dots''.
It should say: ``\dots selection of $f\in\mathscr{H}$ is based\dots''.

\myitem{P 37, Eqs. (3.1) and (3.2)}
\mbox{}\newline
It is NOT advisable to put 2 equation numbers for this only one function!

I guess you now know that the $V$ in (3.1) is wrong.

Moreover, when you write ``$\big(f(\bfx),y\big)\mapsto V\big(f(\bfx),y\big)$''
in (3.2), you are saying that the {\em domain\/} of your function is
$f(X)\times Y$ and not $X\times Y$, where $f(X)\subseteq Y$ but it may
be very well possible that $f(X)\neq Y$ ($f$ could be not {\em onto\/}).

Thus, in $(3.2)$ you are tacitly assuming that the function is defined on
$f(X)\times Y$, where $f(x)\subseteq Y$.

For equation (3.2) to be correct you should write:
\begin{equation}\label{20170819:1}
V:f(X)\times Y\to\R_0^+\,,\qquad
\big(f(\bfx),y\big)\mapsto V\big(f(\bfx),y\big)\,.
\end{equation}

Note the difference between \verb|\to|: $\to$\,, 
and \verb|\mapsto|: $\mapsto$\,.

Note that $\R^+:=\big\{ x\in\R\mid x>0 \big\}$ and
$\R_0^+:=\big\{ x\in\R\mid x\geq0 \big\}$.
$V(f(x),y)$ can be 0!
Hence, instead of $\R^+$ you should write $\R_0^+$.

Here is my suggestion for (3.1), (3.2):
\begin{equation}\label{20170819:2}
V(f(\cdot),\cdot):X\times Y\to\R_0^+\,,\qquad
(\bfx,y)\mapsto V\big(f(\bfx),y\big)
\end{equation}
Note that, although the {\em values\/} of the functions (\ref{20170819:1})
and (\ref{20170819:2}) are the same, namely $V\big(f(\bfx),y\big)$, the
{\em functions\/} $V$ and $V(f(\cdot),\cdot)$ are {\em very\/} different.

\myitem{P 37, L -6}
Formula\quad
$\displaystyle\widehat{f}
=\underset{f\in\mathscr{H}}{\text{arg min\ }} R_{\text{emp}}[f]$\,.

Don't use \verb|\hat| because the hat is too small: $\hat{f}$ !\quad
Use \verb|\widehat| instead: $\widehat{f}$.

You are using \verb|\mathcal{H}|: $\mathcal{H}$.
I like more \verb|\mathscr{H}|: $\mathscr{H}$.
But this is a matter of taste!

\myitem{P 37, formulae (3.3) and (3.6)}
Formula (3.6) tacitly assumes that the distribution $p$ is actually constant
although in L. -9 you wrote: 
``Since the true distribution is {\em unknown\/}...''

There is a contradiction here! Please clarify!

The Riemann sum approximmation of the integral in (3.3) is:
$$
\sum_{i=1}^n V\big( f(\bfx_i,y_i \big)\,p(\bfx_i,y_i)\,\frac1n\,,
$$
which is different fron your formula (3.6).
This formula would coincides with your formula (3.6) in the case
$$
p(\bfx_i,y_i)=1\quad \text{for all}\ i=1,\dots,n\,.
$$
You are tacitly assuming that your data set here is
$S=\big\{ (\bfx_i,y_i)\mid i=1,\dots,n\big\}\in\mathscr{T}$ and
that $p(\bfx,y)=0$ for all $(\bfx,y)\in X\times Y\setminus S$.
Thus everything would be OK for this particular distribution $p$.

\myitem{P 37, formula (3.7)}
The right hand side of formula (3.7) is wrong:
it cannot contain variables (like $\bfx\in X$ or $y\in Y$) because these
would be ``destroyed'' by the operator $E$, producing numerical values.
This operator $E$ acts on {\em functions\/} like $f$ or $V(f(\cdot),\cdot)$.

This formula should be as follows:
$$
\frac1n\,\sum_{i=1}^n V\big( f(\bfx_i,y_i \big)
\overset{n\to\infty}{\longrightarrow} E\big[ V(f(\cdot),\cdot)\big]\,.
$$

\myitem{P 38, Complexity \dots of the hypothesis class + Fig. 3.1}
\mbox{}\newline
It seems that in the left diagram in Fig. 3.1 the data set is
$S=\{z_1,z_2,z_3\}$.
I guess that the possible ``dichotomies'' are the following $3$: 
$D_1=\big\{ \{z_1\}, \{z_2,z_3\} \big\}$,
$D_2=\big\{ \{z_2\}, \{z_1,z_3\} \big\}$,
$D_3=\big\{ \{z_3\}, \{z_1,z_2\} \big\}$.
There are no other.
The two components of each one of the 3 dichotomies $D_1, D_2,D_3$ can
be separated (i.e., shattered) by one line in the {\em plane\/}.

On the right diagram of Fig. 3.1 the data set woul be $S=\{z_1,z_2,z_3,z_4\}$.
The possible dichotomies are then the following $7$:
$D_1=\big\{ \{z_1\}, \{z_2,z_3,z_4\} \big\}$,
$D_2=\big\{ \{z_2\}, \{z_1,z_3,z_4\} \big\}$,
$D_3=\big\{ \{z_3\}, \{z_1,z_2,z_4\} \big\}$.
$D_4=\big\{ \{z_4\}, \{z_1,z_2,z_4\} \big\}$.
$D_5=\big\{ \{z_1,z_2\}, \{z_3,z_4\} \big\}$,
$D_6=\big\{ \{z_1,z_3\}, \{z_2,z_4\} \big\}$,
$D_7=\big\{ \{z_1,z_4\}, \{z_2,z_3\} \big\}$,
If the 4 data $z_1,z_2,z_3,z_4$ are arranged as the vertices of a square
with sides $z_1z_2$, $z_2z_3$, $z_3z_4$, $z_4z_1$ on a plane, then it is
impossible to separate the dichothomy
$D_6=\big\{ \{z_1,z_3\}, \{z_2,z_4\} \big\}$
by means of a straight line on the plane.

In the left side of Fig. 3.1 you say that you shatter {\em the three vectors\/}
by {\em lines\/} in a plane.
You should say that what you separate are the 3 {\em dichotomies\/}.
Dichotomy is the crucial concept here!

In the right side of Fig. 3.1 yo say that is is impossible to shatter appart
the 4 vectors by means of a {\em plane\/}.
This is misleading.
You should talk of {\em straight lines\/}

Both diagrams, left and right, are planar, 2-dimensional diagrams.
So you cannot talk about planes on a plane diagram.

The names $z^1$ and $z^2$ of the coordinate axis, are misleading.
It is better to change them by $x$ and $y$, so that the vectors are
2-dimensional: $z_k=(x_k,y_k)$,  $k=1,2,3,4$.
2-dimensional vectors are enough to explain the concept.

Your definition of the VC-dimension is very obscure.
Compare with the following transcription from Wikipedia
(\texttt{https://en.wikipedia.org/wiki/VC\_dimension}),
where the classification model $f$ is a straight line:

\vspace{2ex}
\red{{\em
``$f$ is a straight line as a classification model on points in a
two-dimensional plane (this is the model used by a perceptron). 
The line should separate positive data points from negative data points. 
There exist sets of 3 points that can indeed be shattered using this
model (any 3 points that are not collinear can be shattered). 
However, no set of 4 points can be shattered: by Radon's theorem,
any four points can be partitioned into two subsets with intersecting
convex hulls, so it is not possible to separate one of these two subsets
from the other. 
Thus, the VC-dimension of this particular classifier is 3.''
}}

\vspace{3ex}
This is a clear description in order to understand the concept of
VC-dimension.
Don't forget the references!
But, of course, in a PhD thesis (or in a paper) it is NOT vey wise to
quote Wikipedia.
You must go to the original sources.

A key point is Radon's theorem in discrete geometry
($\text{co}(A)$ denotes the convex hull of $A$):

\begin{theorem}[J. Radon]
Let $S\subset\R^d$ be a subset containing at least $d+2$ elements. 
Then there exist {\bf disjoint} subsets $A\subset S$ and $B\subset S$
such that
$$
\text{co}(A)\cap\text{co}(B)\neq\emptyset
$$
\end{theorem}
\noindent
Ref.: B.B. Peterson, {\em ``The Geometry of Radon's Theorem''\/}, The
American Mathematical Monthly, Vol. 9, No. 9 (Nov., 1972), 949-963.

\vspace{2ex}
Radon's theorem implies that in $\R^d$ any set of $d+2$ or more points
(vectors!) has at least one dichotomy which cannot be separated (shattered)
by any plane in $\R^d$.

Thus the VC-dimension in $\R^d$ cannot be greater than $d+1$.

\end{description}

%% END OF FILE