### 2017-09-23 src/0_Abstract.tex
Conclusions first part

\chapter*{Abstract}

\emph{Financial time series are known for their non-stationary behaviour and
this has motivated its study using different techniques. 
One relevant observation is that sometimes time series exhibit some stationary
linear combinations. When this happens, it is said that those time series are
cointegrated. Cointegration has been then one of the main features studied.
Vector error correction model (VECM) is an econometric model which characterises
the joint dynamic behaviour of a set of cointegrated variables in terms of
forces pulling towards equilibrium.}

\emph{Cointegration relationships are found on different time series
frequencies. Cointegration in low frequency time series is motivated by a
long-run equilibrium relationship between economic forces, whereas cointegration
in high frequency data has its foundation in statistical arbitrage theory.
However, the use of cointegration models such as VECM is mainly limited by
computationally expensive routines.}

\emph{In this thesis, financial time series features are studied and different
algorithms were developed in order to optimise parameters and increase
performance.  In particular, an online algorithm based on VECM which optimises
how model parameters are obtained and reduces execution times is proposed. This
is achieved considering only a sliding window of the last historical data and
using machine learning techniques to solve the model. Moreover, the long-run
relationship between the time series is used in order to make optimisations and
obtain improved execution times.  Due to the large amount of financial data
available and the need of quick response, the algorithm presented was optimised
using high performance computing. The experiments were tested using foreign
exchange rates.  Results show that cointegration and high performance computing
allow to obtain models with better performance accuracy and reduced execution
times.}

\vspace{0.8cm}

\emph{\textbf{Keywords:} Computational finance  - Cointegration  - Time Series - 
Online algorithms  - Regression}