### 2017-10-15 src/firmas.tex
fixing table captions

\thispagestyle{empty}
\cleardoublepage%
\bigskip
\bigskip
\bigskip
\bigskip
\bigskip
\vspace{1cm}
\begin{center}
\includegraphics[scale=0.6]{fig/logo_usm}\\
\cambiaFont{ppl}{\Large \textsc{Examining Committee}}\\
\bigskip
\end{center}
\vspace{2cm}
\begin{tabular}[c]{lc}
{\textsc Prof. Dr. Luis Salinas} & \begin{picture}(150,5) \line(1,0){150} \end{picture}\\
{Universidad T\'ecnica Federico Santa Mar\'ia} & {Advisor}\\
Chile & \\
& \\
& \\
{\textsc Dr. Claudio Torres} &  \begin{picture}(150,5) \line(1,0){150} \end{picture}\\
{Universidad T\'ecnica Federico Santa Mar\'ia} & {Co-advisor} \\
Chile & \\

& \\
& \\
{\textsc Dr. Werner Kristjanpoller} &  \begin{picture}(150,5) \line(1,0){150} \end{picture}\\
{Universidad T\'ecnica Federico Santa Mar\'ia} & {Co-advisor} \\
Chile & \\
& \\
& \\
{Prof. Dr. Christian de Peretti} & \begin{picture}(150,5) \line(1,0){150} \end{picture}\\
{Ecole Centrale de Lyon} & {External Peer Reviewer}\\
{France} & \\
& \\
& \\
{Prof. Dr. Wilfredo Palma} & \begin{picture}(150,5) \line(1,0){150} \end{picture}\\
{Pontificia Universidad Cat\'olica de Chile} & {External Peer Reviewer}\\
{Chile} & \\
& \\
& \\
{\textsc Prof. Dr. Hern\'an Astudillo} &  \begin{picture}(150,5) \line(1,0){150} \end{picture}\\
{Universidad T\'ecnica Federico Santa Mar\'ia} & { PhD program Director} \\
Chile & \\
%& \\
%& \\
%{\textsc Prof. Dr. Carlos Castro} &  \begin{picture}(150,5) \line(1,0){150} \end{picture}\\
%{Universidad T\'ecnica Federico Santa Mar\'ia} & {Committee Member} \\
%Chile & \\
\end{tabular}

\cleardoublepage%