### 2017-12-18 articles-conferences/jcc2012/JCC2012/src/AAR.tex
Adding all before the exam

\section{The Aggregating Algorithm for Regression}

The AAR, proposed by~\cite{vovk2001}, is an application of the
aggregating algorithm to the problem of regression.

The prediction formula for AAR is given by
equation~(\ref{eq:predAAR}):

\begin{equation}
\label{eq:predAAR}
f(\mathbf{x}_{m+1}) = \displaystyle \big( \sum_{t=1}^m y_t \mathbf{x}_t \big )^\intercal 
  \big ( \sum_{t=1}^{m+1} \mathbf{x}_t \mathbf{x}_t  ^\intercal + \gamma
\mathbb{I})\big )^{-1}  \mathbf{x}_{m+1} \, ,
\end{equation}

\noindent which is very similar to the RR method except because AAR includes
information of the new input $\mathbf{x}_{m+1}$.  This means that AAR
updates matrix $A$ with the new input $\mathbf{x}_{m+1}$ before the
prediction $f(\mathbf{x}_{m+1})$ is made.

Algorithm~\ref{alg:AAR} shows this procedure based on
equation~(\ref{eq:predAAR}). This algorithm can be obtained by swapping lines 4 and 5 from
algorithm~\ref{alg:RR}.

\begin{algorithm}[ht]
\begin{algorithmic}[1]
\REQUIRE $\,$ \\
$\{\mathbf{x}_1,\dots,\mathbf{x}_m \}$: $m$ input vectors \\
$\{y_1,\dots,y_m \}$: $m$ targets \\
\ENSURE  $\,$ \\
$\{f(\mathbf{x}_1),\dots,f(\mathbf{x}_m) \}$: model predictions \\
\STATE Initialize $\mathbf{A}=\gamma \mathbb{I}$
and $\mathbf{b}=0$
\FOR { $t = 1$ to $m$ }
   	\STATE read new $\mathbf{x}_t$
	\STATE $\mathbf{A} = \mathbf{A} + \mathbf{x}_t \mathbf{x}_t^\intercal$
	\STATE output prediction $f(\mathbf{x}_t) =  \mathbf{b}^\intercal \mathbf{A}^{-1}\mathbf{x}_t$
   	\STATE Read new $y_t$
    	\STATE $\mathbf{b} = \mathbf{b} + y_t \mathbf{x}_t$
\ENDFOR

\end{algorithmic}
\caption{{\em The aggregating algorithm for regression}}
\label{alg:AAR}
\end{algorithm}