### 2017-12-18 articles-conferences/jcc2012/JCC2012/src/experiments.tex

\section{Experimental Results}

In order to measure performance of our algorithm, we used daily stock
returns of 45 stocks from the technology sector dated from the 1st of
January 2000 to the 1st of July 2012.

\begin{figure}[h]
  \centering
  \subfigure[AAR v/s the target for SPY]{
  \includegraphics[width=0.4\textwidth]{img/AAR}
  %\caption{AAR v/s the target for SPY}
  \label{fig:data}
  }
  \subfigure[RR v/s the target for SPY]{
  \includegraphics[width=0.4\textwidth]{img/RR}
  %\caption{AAR v/s the target for SPY}
  \label{fig:data2}
  }
  \subfigure[SLAAR v/s the target for SPY]{
  \includegraphics[width=0.4\textwidth]{img/SLAAR}
  %\caption{AAR v/s the target for SPY}
  \label{fig:data3}
  }
\end{figure}

Returns $\{x_t\}_{t=1}^{m-1}$ were defined in based on stock prices
$\{{p_t}\}_{t=1}^m$ and are related as:

\begin{equation*}
x_t = \frac{p_{t+1}-p_t}{p_t} \, .
\end{equation*}

The objective is to build a model of the returns of stock $k$ based on
returns of other stocks of the same financial sector.

Therefore, every input vector $\mathbf{x}_t$ will have 44 stock
returns at time $t$ without considering information of stock $k$. The
target vector $\mathbf{y}$ will be the stocks returns of stock $k$.

We compared RR, AAR and our proposal called sliding windows AAR
(SLAAR) and the results are shown in table~\ref{tab:results}. The MSE
was shown in bold when our method was better than RR and AAR. The
table shows that our method outperforms in 28 of 45 stocks.

\begin{table}[h!]
    \centering
    \footnotesize
    \begin{tabular}{*{5}{r}}
        \toprule
          & AAR & RR & SLAAR& Best L \\
         \midrule
Stock  &  AAR error  &   RR error  &   SLAAR error  &   Best L  \\
IBM  &  0.230307  &  0.231317  &  \textbf{0.228963}  &  70   \\
MMM  &  0.199237  &  0.197172  &  \textbf{ 0.195550} &  990\\
CVX  &  0.155454  &  0.152070  &  0.155150  &  750\\
UTX  &  0.260980  &  0.268351  &  0.261140  &  340\\
CAT  &  0.223338  &  0.222240  &  \textbf{ 0.210500} &  360\\
MCD  &  0.268211  &  0.266997  &  0.267182  &  990\\
BA  &  0.292699  &  0.295623  &  \textbf{ 0.292180} &  150\\
XOM  &  0.160376  &  0.157125  &  0.159074  &  780\\
JNJ  &  0.213276  &  0.214747  &  \textbf{ 0.212601}  &  680\\
PG  &  0.494937  &  0.493036  &  0.495395  &  950\\
KO  &  0.272203  &  0.274337  &  \textbf{ 0.270888} &  870\\
WMT  &  0.272945  &  0.269190  &  \textbf{ 0.254820}  &  630\\
HPQ  &  0.276595  &  0.275459  &  0.280236  &  980\\
AXP  &  0.204162  &  0.211216  &  \textbf{ 0.195888}  &  230\\
DD  &  0.204220  &  0.202591  &  0.203177  &  330\\
JPM  &  0.238405  &  0.255824  &  \textbf{ 0.204777}  &  290\\
MRK  &  0.214294  &  0.211708  &  \textbf{ 0.196368}  &  850\\
DIS  &  0.324324  &  0.326105  &  \textbf{ 0.297895}  &  290\\
VZ  &  0.203936  &  0.206390  &  \textbf{ 0.196078}  &  490\\
HD  &  0.236339  &  0.237688  &  0.242606  &  990\\
T  &  0.235940  &  0.234002  &  \textbf{ 0.230230}  &  570\\
MSFT  &  0.332521  &  0.331694  &  0.334283  &  710\\
CSCO  &  0.227675  &  0.228196  &  0.233057  &  990\\
INTC  &  0.262545  &  0.264391  &  0.271204  &  1000\\
GE  &  0.168935  &  0.168258  &  0.168924  &  1000\\
PFE  &  0.200335  &  0.199163  &  0.204552  &  1000\\
BAC  &  0.299956  &  0.333396  &  \textbf{ 0.200134}  &  530\\
AA  &  0.249913  &  0.250263  &  \textbf{ 0.246067}  &  1000\\
XLF  &  0.300486  &  0.299604  &  0.301322  &  830\\
EWH  &  0.186584  &  0.188475  &  \textbf{ 0.180674}  &  260\\
EWG  &  0.156459  &  0.158946  &  \textbf{ 0.152092}  &  470\\
EWA  &  0.231424  &  0.233347  &  \textbf{ 0.224692}  &  960\\
XLV  &  0.140344  &  0.139772  &  \textbf{ 0.118340}  &  610\\
XLI  &  0.089607  &  0.087342  &  \textbf{ 0.086304}  &  450\\
XLU  &  0.107657  &  0.106929  &  \textbf{ 0.091362}  &  550\\
XLY  &  0.103068  &  0.102359  &  \textbf{ 0.095318}  &  590\\
XLB  &  0.105090  &  0.102791  &  \textbf{ 0.099280}  &  940\\
XLE  &  0.118998  &  0.117319  &  \textbf{ 0.111371} &  1000\\
SPY  &  0.073800  &  0.071908  &  \textbf{ 0.067655} &  230\\
EWS  &  0.223738  &  0.223721  &  0.228166  &  990\\
EWC  &  0.172954  &  0.174107  &  \textbf{ 0.164855} &  330\\
EWU  &  0.146232  &  0.145031  &  0.146491  &  640\\
EWW  &  0.205101  &  0.203012  &  \textbf{ 0.199228}  &  80\\
EWM  &  0.249645  &  0.248694  &  0.257393  &  80\\
XLK  &  0.098086  &  0.098599  &  \textbf{ 0.097944}  &  440  \\

        \bottomrule
    \end{tabular}
    \caption{MSE and best L}
    \label{tab:results}
\end{table}

%\begin{figure}[h!]
%  \centering
%    \includegraphics[width=0.55\textwidth]{img/results}
%  \caption{Comparison AAR, RR and SLAAR for SPY}
%  \label{fig:data}
%\end{figure}

%\begin{figure}[h!]
%  \centering
%    \includegraphics[width=0.45\textwidth]{img/RR}
%  \caption{RR v/s the target for SPY}
%  \label{fig:data2}
%\end{figure}
%
%
%\begin{figure}[h!]
%  \centering
%    \includegraphics[width=0.45\textwidth]{img/SLAAR}
%  \caption{SLAAR v/s the target for SPY}
%  \label{fig:data3}
%\end{figure}

The error was calculated considering the {\em mean squared error}
(MSE) using the last 70 predictions made.

Table~\ref{tab:results} also shows that the smaller error in our
method was for the SPY stock, which is an expected result considering
that the SPY value is constructed based on the main stocks of the
markets.

A graphical comparison between AAR, RR and SLAAR with the target for
stock SPY is shown in figures~\ref{fig:data},\ref{fig:data2} and
\ref{fig:data3}. It shows how AAR and our method fit better than RR to
the target and also that our method is slightly better compared with
AAR. It's also possible to notice that when returns change drastically
none of the methods fit satisfactorily.

\section{Discussion and Conclusion}

The existing approaches to solve regression problems are well known,
however they are not used in an online context because they involve
many calculations and in order to make a prediction, they consider all
the data available. However, in some cases, only a portion of data is
needed or available to make a prediction and sometimes the accuracy
could be improved considering less data.

In this article, this historical dependence of data is studied and the
results show that in some cases, if we use a sliding window of data
instead of all the data, the MSE of the predictions is minimized.
It's possible to check that there are some stocks which depend more on
the historical data than others and this is valuable information for
future algorithms.

On the other hand, RR and AAR calculates an inverse matrix at every
step, which is computationally very expensive. In our proposal, this
inverse matrix calculation is avoided and therefore it's more suitable
to be used in an online context.