### 2017-12-18 articles-conferences/jcc2012/JCC2012/src/intro.tex

\section{Introduction}
The learning from examples problem is an ill-posed problem which
admits an infinite number of solutions.  In order to restrict the
space of admissible solutions, the regression problem is usually
formulated in terms of regularization theory~\cite{girosiETal1995} as
an optimization problem, which minimizes the functional:

\begin{equation}
\label{eq:problem} 
J(\mathbf{w}) = \sum_{t=1}^m (y_t - f(\mathbf{x}_t))^2 + \gamma R(\mathbf{w})
, \qquad f \in \mathcal{H} \, ,
\end{equation}

\noindent where $m$ is the number of samples $(\mathbf{x}_t ,y_t)$ with
$\mathbf{x}_t \in \R^l$, $l$ correspond to the number of features, $y_t
\in \R$ is the target, $R(\mathbf{w}) = ||w||^2$ where $||.||^2_K$ is a norm in a
{\em reproducing kernel hilbert space} $\mathcal{H}$ defined by the
positive definite form $K$, and $\gamma$ is a regularization
parameter~\cite{evgeniouETal2000}.
%Qué norma debe usarse? falta definir K
When the hypothesis space is reduced, the risk of overfitting the training data
decreases and therefore leading to better generalization capability. 
%aclarar porqué se reduce el espacio de hipótesis
{\em Ridge regression} (RR) is a batch method generally used to solve this
problem, which is a generalization of least squares method (LS).

%RR is a batch method and it is not used in an online context mainly because
%it uses large amount of data and the number of operations increases
%with data size.

However, online algorithms are more attractive than batch algorithms because
their simplicity and ability to manage large data sets, this is why
they are popular in financial applications.

There are several popular online methods such as
perceptron~\cite{rosenblatt58},
passive-aggressive~\cite{crammerETall2006}, stochastic gradient
descent~\cite{zhang2004}, aggregating algorithm~\cite{vovk2001} and
the second order perceptron~\cite{cesa-bianchi2005}.
In~\cite{cesa-bianchi2006} it is provided an in-deph analysis of
online learning.

The {\em aggregating algorithm for regression} (AAR) method formulates
a recursive formulation of RR in an online way. AAR
consider all data to make a prediction, but in highly variant
scenarios the old data could be useless.

In this paper, we propose an online method based on the idea presented
by~\cite{vovk2001} considering in our case a single sliding window of
the most recent data. This proposal also reduces the number of
operations at every step of the algorithm by expressing the inverse
matrix in an iterative form. Our algorithm is later tested with
financial data from stock market.